% Options for packages loaded elsewhere
% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  11pt,
  letterpaper,
  numbers=noendperiod,
  parskip=half]{scrartcl}
\usepackage{xcolor}
\usepackage[margin=1in,heightrounded]{geometry}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{3}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{libertinus}
\ifPDFTeX\else
  % xetex/luatex font selection
  \setmainfont{Libertinus Serif}
  \setsansfont{Libertinus Sans}
  \setmonofont{JetBrains Mono}[Scale=0.88]
  \setmathfont{Libertinus Math}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\newenvironment{Shaded}{}{}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.33,0.33}{\textbf{#1}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.42,0.45,0.49}{#1}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.84,0.23,0.29}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.36,0.77}{#1}}
\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.84,0.23,0.29}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.01,0.18,0.38}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.42,0.45,0.49}{#1}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.42,0.45,0.49}{#1}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.36,0.77}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.84,0.23,0.29}{#1}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.84,0.23,0.29}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.36,0.77}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.42,0.45,0.49}{#1}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.33,0.33}{\underline{#1}}}
\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.84,0.23,0.29}{\textbf{#1}}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.36,0.77}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.44,0.26,0.76}{#1}}
\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.01,0.18,0.38}{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.42,0.45,0.49}{#1}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.84,0.23,0.29}{#1}}
\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.14,0.16,0.18}{#1}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.14,0.16,0.18}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.44,0.26,0.76}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.84,0.23,0.29}{#1}}
\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.42,0.45,0.49}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.36,0.77}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.01,0.18,0.38}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.01,0.18,0.38}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.89,0.38,0.04}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.01,0.18,0.38}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{1.00,0.33,0.33}{#1}}

\usepackage{longtable,booktabs,array}
\newcounter{none} % for unnumbered tables
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{pgfplots}
\usepgfplotslibrary{groupplots}
\usepackage{tabularx}
\usepackage{makecell}
\usepackage{adjustbox}
\usepackage{siunitx}
\usepackage{colortbl}
\usepackage{enumitem}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother


% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}



\setlength{\emergencystretch}{3em} % prevent overfull lines

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}



 


\usepackage{microtype}
\usepackage{hyperref}
\hypersetup{
  pdftitle={Context Routing for AI Coding Agents},
  pdfauthor={Gunther Brunner},
  pdfsubject={Context allocation policies for AI coding agents},
  pdfkeywords={AI agents, context routing, context engineering, retrieval, vendoring, MCP, agent skills}
}
\usepackage{titling}
\pretitle{\begin{center}\fontsize{22pt}{26pt}\selectfont\bfseries}
\posttitle{\par\end{center}\vskip 0.5em}
\preauthor{\begin{center}\large}
\postauthor{\end{center}}
\predate{\begin{center}\large}
\postdate{\par\end{center}\vskip 1em}
\usepackage[automark,headsepline,footsepline]{scrlayer-scrpage}
\clearpairofpagestyles
\ihead{\small Context Routing for AI Coding Agents}
\ifoot{\small Brunner, 2026}
\ofoot{\pagemark}
\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\@ifpackageloaded{fontawesome5}{}{\usepackage{fontawesome5}}
\definecolor{paper-callout-color}{HTML}{909090}
\definecolor{paper-callout-note-color}{HTML}{0758E5}
\definecolor{paper-callout-important-color}{HTML}{CC1914}
\definecolor{paper-callout-warning-color}{HTML}{EB9113}
\definecolor{paper-callout-tip-color}{HTML}{00A047}
\definecolor{paper-callout-caution-color}{HTML}{FC5300}
\definecolor{paper-callout-color-frame}{HTML}{acacac}
\definecolor{paper-callout-note-color-frame}{HTML}{4582ec}
\definecolor{paper-callout-important-color-frame}{HTML}{d9534f}
\definecolor{paper-callout-warning-color-frame}{HTML}{f0ad4e}
\definecolor{paper-callout-tip-color-frame}{HTML}{02b875}
\definecolor{paper-callout-caution-color-frame}{HTML}{fd7e14}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{HTML}{4a90d9}}{}
\makeatother
\makeatletter
\@ifundefined{codebgcolor}{\definecolor{codebgcolor}{HTML}{f5f5f5}}{}
\makeatother
\makeatletter
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, breakable, colback={codebgcolor}, enhanced, frame hidden, sharp corners]}{\end{tcolorbox}}\fi
\makeatother
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Context Engineering for AI Coding Agents --- A Field Guide to Strategies, Tools, and Fit},
  pdfauthor={Gunther Brunner},
  pdfkeywords={AI coding agents, context engineering, retrieval-augmented
code generation, repository-level code retrieval, vendoring,
Model Context Protocol, agent skills, software engineering tools,
practitioner field guide},
  colorlinks=true,
  linkcolor={RoyalBlue},
  filecolor={Maroon},
  citecolor={RoyalBlue},
  urlcolor={RoyalBlue},
  pdfcreator={LaTeX}}

\input{figures.tex}


\title{Context Engineering for AI Coding Agents}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{A Field Guide to Strategies, Tools, and Fit}
\author{Gunther Brunner}
\date{2026-05-15}
\begin{document}
\maketitle
\begin{abstract}
AI coding agents do real work now --- patching production repositories,
migrating frameworks, fixing failing tests, answering questions about
large codebases --- and their success rate is bounded by something less
fashionable than raw model capability: \emph{which context they have at
hand and how it got there.} A million-token window does not guarantee
that the right tokens are present, attended to, version-correct, or
trustworthy.

This is a practitioner's field guide to the answer space. It does three
things:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Catalogs the seven ways context flows into an agent today --- vendoring,
  lazy fetch, repo packing, prose rules, pre-built indices, code overlays,
  and tool-mediated agentic search --- with the mechanism, the winning
  conditions, the failure modes, and the cost profile of each.
\item
  Surveys the May 2026 tooling landscape grouped by which strategy each
  tool implements: 80+ products spanning IDE-native context engines, MCP
  servers, embeddings/RAG, code search, repo packers, multi-repo systems,
  and adjacent safety/workflow tools.
\item
  Offers a fit guide --- a decision matrix and a set of worked scenarios
  --- that connects task class, scale, and constraints to a concrete
  combination of strategies and tools that suits the job.
\end{enumerate}

Empirical evidence (LongCodeBench, CodeRAG-Bench, RULER, structural
retrieval studies, distractor sensitivity work) sits underneath the guide
as supporting material: it explains \emph{why} the recommendations break
the way they do, without becoming the paper's central argument. Standards
and interoperability (MCP, AGENTS.md, Skills, llms.txt) are surveyed
briefly because they determine which combinations are portable across
agents.

The guide is shaped by work on \emph{ingraft}, an open-source CLI for
vendoring upstream source into agent-ready repositories, but ingraft is
treated here as one slot in the landscape rather than the protagonist.
The reader's question is the protagonist: ``what should I use for this
job?''
\end{abstract}

\renewcommand*\contentsname{Table of contents}
{
\hypersetup{linkcolor=black}
\setcounter{tocdepth}{3}
\tableofcontents
}

\section{Introduction}\label{sec-intro}

AI coding agents are now useful enough that their failures matter.
Developers ask them to patch production repositories, migrate
frameworks, inspect failing tests, answer questions about large
codebases, and integrate unfamiliar libraries. In these settings, the
bottleneck is rarely ``no relevant information exists.'' The bottleneck
is context allocation: which information should enter the agent's
working set, in which representation, through which trust boundary, at
what point in the loop, and under what budget.

\figureContextAtGlance

This guide treats that question as a practical problem with practical
answers, not a research thesis. Seven distinct ways of moving context
into an agent are in active use today, dozens of products implement each
one, and most real tasks need a \emph{combination} of two or three. The
goal of the guide is to make those choices legible to the person who
has to make them.

\subsection{What this guide covers}\label{sec-guide-scope}

The body answers a reader's questions in order:

\begin{itemize}
\tightlist
\item
  \textbf{\S\ref{sec-strategies}, The strategies} --- the seven ways
  context flows into an agent today. For each: how it works, where it
  wins, where it loses, what it costs.
\item
  \textbf{\S\ref{sec-tools}, The tools} --- the May 2026 product catalog
  grouped by which strategy each implements.
\item
  \textbf{\S\ref{sec-fit}, The fit guide} --- a decision matrix and
  seven worked scenarios that connect task class, scale, and
  constraints to a concrete combination of strategies and tools.
\item
  \textbf{\S\ref{sec-evidence}, The evidence} --- the research papers
  and benchmarks that explain \emph{why} the recommendations break the
  way they do.
\item
  \textbf{\S\ref{sec-routing}, How to think about routing} --- a
  framework (placement, representation, selector, timing, constraints)
  for reasoning about a route when the decision matrix does not cover
  the case.
\item
  \textbf{\S\ref{sec-standards}, Standards and interop} --- the
  ecosystem-wide protocols (MCP, AGENTS.md, Skills, llms.txt) that
  determine which combinations are portable across agents.
\item
  \textbf{\S\ref{sec-ingraft}, Where ingraft fits} --- one slot in the
  vendoring strategy.
\end{itemize}

The guide is a dated snapshot. Most tool and standards claims were
checked against public documentation, repositories, specifications, or
vendor posts available during March--May 2026. Tool counts and product
features age fast.

The immediate product trigger was Maxwell Brown's Effect post on using
git subtrees for coding-agent context (Brown 2026): ``coding agents
rely on patterns and examples, not descriptions; give your agent real
library code to learn from.'' That claim is intentionally narrower than
this guide. It does not settle the general problem; it points at it: if
subtree is the right route for Effect, what decides when subtree is
right, when docs are better, and when neither should be loaded? The
field-guide answer is: it depends on the task, the scale, and the
constraints --- and the rest of the document is the dependency table.

\subsection{Vocabulary}\label{sec-vocabulary}

Six terms recur throughout. They are defined once here.

\begin{itemize}
\item
  \textbf{Context} --- any information made available to an agent during
  a task: prompt tokens, files it can read, tool results, memory,
  indexes, skills, instructions, logs, external service responses.
\item
  \textbf{Context object} --- a knowledge unit that might be relevant: a
  package, file, repository, issue, stack trace, lockfile entry, API
  document, design note, generated summary, skill.
\item
  \textbf{Strategy} --- a way of exposing a context object to the agent.
  The seven strategies in \S\ref{sec-strategies} are the catalog of
  options.
\item
  \textbf{Bundle} --- the concrete material exposed to the agent for a
  given task: selected files, docs, snippets, index hits, rules, tool
  affordances.
\item
  \textbf{Authority} --- which source should win for a specific
  question: the lockfile for installed version, official docs for public
  API intent, source and tests for behavior, local code for project
  convention, human instructions for constraints.
\item
  \textbf{Provenance} --- where a bundle came from, which version it
  represents, when it was fetched, how trustworthy it is, and what it
  is intended to be used for.
\end{itemize}

A seventh term, \textbf{context router}, appears in \S\ref{sec-routing}
as the name for the explicit policy that selects among strategies. It
is one organizing lens, not a prescription. The guide works whether or
not you adopt the term.

%====================================================================
%  CUT: Method and Scope, Definitions, Utility Model, Policy Signals.
%  The essential vocabulary is absorbed into \S\ref{sec-vocabulary}.
%  The utility model and policy signals tables are repositioned in
%  \S\ref{sec-routing} where they are operationally relevant.
%====================================================================
\iffalse
\section{Method and Scope}\label{sec-method}

This paper is a dated synthesis, not an exhaustive census. The review
covers materials visible through May 13, 2026, with most tool and
standards claims checked against public documentation, repositories,
specifications, or vendor posts available during March-May 2026.

\subsection{Search and inclusion criteria}\label{sec-method-search}

Sources searched included arXiv, ACL Anthology, ACM/IEEE proceedings
where available, OpenReview, GitHub, vendor documentation, engineering
blogs, standards websites, tool registries, Hacker News, practitioner
podcasts, and industry surveys.

Papers were included when they addressed at least one of these
questions:

\begin{itemize}
\tightlist
\item
  long-context evaluation for language models or coding models;
\item
  repository-level code generation, completion, repair, or localization;
\item
  retrieval-augmented code generation;
\item
  semantic, symbolic, graph, or AST-based code retrieval;
\item
  agent scaffolding, context management, compression, or distractor
  behavior.
\end{itemize}

Tools were included when they affected how coding agents obtain
repository, library, documentation, organizational, or procedural
context. The appendix is a representative May 2026 snapshot, not a
complete inventory. Tool counts and product features are volatile.

Excluded materials include pure model benchmarks with no context
intervention, general-purpose RAG systems with no code or agent
relevance, and abandoned tools unless they were historically
influential.

\subsection{Evidence hierarchy}\label{sec-evidence-hierarchy}

The paper uses heterogeneous evidence. To keep claims calibrated,
Table~\ref{tbl-evidence} separates the role of each evidence class.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}@{}}
\caption{Evidence classes and their role in the
argument.}\label{tbl-evidence}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Evidence class
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Examples
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
How it is used
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Evidence class
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Examples
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
How it is used
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Peer-reviewed empirical papers & Lost in the Middle, RULER, RepoBench,
SWE-bench, GraphCoder & Mechanisms and benchmark results with
comparatively strong evidential weight \\
Preprints and technical reports & LongCodeBench, cAST, Context Rot,
newer repo-retrieval papers & Emerging findings, cited with venue
caveats \\
Vendor engineering posts & Anthropic, Cursor, Cognition, Augment,
Sourcegraph & Evidence of deployed design patterns and product
constraints, not independent proof \\
Practitioner essays and podcasts & Effect, Cline, aider, HumanLayer,
Ptacek, Ronacher & Operating experience and design disagreement \\
Tool documentation and standards & MCP, Agent Skills, AGENTS.md,
llms.txt, SCIP & Implementation facts and interoperability surfaces \\
This paper's synthesis & router thesis, policy model, evaluation design
& Interpretation that should be tested, not treated as direct empirical
result \\
\end{longtable}

Two terms are used carefully:

\begin{itemize}
\tightlist
\item
  \textbf{Academic and technical works} means peer-reviewed papers,
  workshop papers, arXiv preprints, benchmark reports, and technical
  reports. It does not imply every cited work is peer-reviewed.
\item
  \textbf{Representative tool snapshot} means a dated inventory of
  visible major tools. It is not complete and will age.
\end{itemize}

Practitioner evidence shows convergence around a shared problem, not
consensus on the solution. The field agrees that context selection
matters; it disagrees on where selection should happen.

\section{Definitions and Problem Formulation}\label{sec-formal}

\subsection{Core definitions}\label{sec-definitions}

\textbf{Context} is any information made available to an agent during a
task. It includes prompt tokens, files the agent can read, tool results,
memory, indexes, skills, instructions, logs, and external service
responses.

\textbf{Context object} is a knowledge unit that might be relevant: a
package, file, repository, issue, stack trace, lockfile entry, API
document, design note, generated summary, or skill.

\textbf{Context bundle} is the concrete material exposed to the agent
for a task: selected files, docs, snippets, index hits, rules,
provenance notes, and tool affordances.

\textbf{Routing strategy} is a way of exposing a context object: vendor
it into the workspace, fetch it lazily, retrieve docs, query a graph
index, pack a repo, activate a skill, or let an agent inspect through
shell tools.

\textbf{Context router} is a policy
\texttt{\textbackslash{}pi(t,\ m,\ r,\ k)\ -\textgreater{}\ s}: given a
task \texttt{t}, model or agent capability \texttt{m}, repository state
\texttt{r}, and knowledge object \texttt{k}, select a routing strategy
\texttt{s}.

\textbf{Source authority} describes which source should win for a
specific question: the lockfile for installed version, official docs for
public API intent, source and tests for behavior, local code for project
convention, and human instructions for constraints.

\textbf{Context provenance} records where a bundle came from, which
version it represents, when it was fetched, how trustworthy it is, and
what it is intended to be used for.

\subsection{Utility model}\label{sec-utility}

A lightweight model is enough to make the tradeoff explicit. For a
coding task \texttt{t}, model or agent \texttt{m}, repository state
\texttt{r}, knowledge object \texttt{k}, and candidate strategy
\texttt{s}, the router wants high expected success and low total cost:

\begin{align*}
U(s \mid t,m,r,k) ={}&
P(\mathrm{success} \mid t,m,r,C_s) \\
&- \lambda_1\mathrm{tokens}(C_s)
- \lambda_2\mathrm{latency}(s)
- \lambda_3\mathrm{privacyRisk}(s) \\
&- \lambda_4\mathrm{staleness}(s)
- \lambda_5\mathrm{distractorRisk}(C_s)
- \lambda_6\mathrm{trustRisk}(s)
\end{align*}

The context router chooses the strategy with the highest expected
utility under observable signals. The exact weights are product and
organization specific, but the terms make clear why ``more context'' is
not the objective.

\figureUtilityTradeoff

The operational principle is \textbf{smallest sufficient context}:

\begin{quote}
Context is sufficient when it lets the agent identify the relevant
authority, produce a correct change, and verify it without introducing
materially misleading distractors.
\end{quote}

Smallest does not mean fewest tokens. It means lowest total context cost
across tokens, latency, privacy, staleness, trust, and distractor risk.
Sometimes the lowest-cost context is a small doc page. Sometimes it is a
large but version-matched source tree.

\subsection{Policy signals}\label{sec-policy-signals}

The router cannot observe expected success directly, so it uses signals.
The signals in Table~\ref{tbl-signals} are enough to turn ``router''
from metaphor into a policy surface.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.5000}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.5000}}@{}}
\caption{Observable signals for a context-routing
policy.}\label{tbl-signals}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Signal
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Why it matters
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Signal
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Why it matters
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Package appears in lockfile & Fixes version-specific context and avoids
latest-doc drift \\
Import frequency & Estimates package centrality and likely value of
deeper source \\
Package source size & Affects vendoring, packing, indexing, and
distractor cost \\
Docs availability and quality & Makes docs-first viable for public API
tasks \\
API volatility & Increases value of fresh docs, changelogs, or source \\
Local modifications & Penalizes stale indexes and remote snapshots \\
Model context window and measured effective length & Bounds what can be
reliably carried \\
Agent tool capability & Determines whether agentic search is viable \\
Privacy policy & Rules out cloud retrieval or remote indexing \\
Task type & Bugfix, migration, greenfield integration, test repair,
security review, or refactor \\
Failure history & Repeated failures justify deeper context or a strategy
switch \\
Repository size and churn & Determines index/search break-even \\
Language ecosystem support & Determines whether AST, LSP, or SCIP
retrieval is available \\
\end{longtable}
\fi
%====================================================================
%  END CUT
%====================================================================

\section{The Strategies}\label{sec-strategies}

Seven distinct strategies move context from somewhere outside the agent
into the agent's working set. Almost every product on the market
implements one or more of them; the difference between products is which
combination they bake in. This chapter is the catalog: for each
strategy, the mechanism, where it wins, where it fails, and the cost
profile.

A reader who only reads this chapter should be able to recognize, given
any tool, which strategy or combination of strategies the tool
implements --- and therefore which jobs the tool was built to do.

\subsection{Strategy 1 --- Vendoring (source becomes part of the
workspace)}\label{sec-strat-vendor}

\textbf{Mechanism.} Upstream source is committed or mounted directly
inside the project, typically under a reference directory like
\texttt{vendor/}, marked read-only for humans and visible to the agent.
Implementations range from raw \texttt{git subtree}, \texttt{git
submodule}, and \texttt{git clone} to wrappers like ingraft, Vendir,
and Carvel that automate update, ignore-pattern injection, and
provenance.

\textbf{Wins when:} a single package is a load-bearing idiom source for
the project (Effect, Inertia.js, complex internal frameworks); the
agent needs concrete patterns and tests, not prose descriptions; the
source is navigable and version-matched; privacy rules out cloud
indexing; offline operation matters.

\textbf{Loses when:} the package is huge, generated, or full of distractors
(Chromium-scale repos); the project will get tens of dependencies vendored
and the repo bloats; the codebase needs public-API \emph{intent} more than
implementation behavior (docs win there); the team needs to upgrade often
and the wrapper does not automate updates.

\textbf{Cost profile.} High upfront setup and ongoing maintenance.
Near-zero per-request latency and per-request token cost (the agent
just reads files). High determinism. Privacy is local-by-default.

\subsection{Strategy 2 --- Lazy fetch (on-demand pull into a side
cache)}\label{sec-strat-fetch}

\textbf{Mechanism.} The agent asks for a package or doc, the tool
resolves it (npm, PyPI, crates.io, etc.), and the result lands in a
side cache (\texttt{\~{}/.opensrc}, \texttt{node\_modules\_cache}). The
cache is read like a local filesystem, then either retained or pruned.

\textbf{Wins when:} most packages are used occasionally and not worth
vendoring; the project's dependency surface is large but shallow; one
or two packages may need deeper inspection during debugging.

\textbf{Loses when:} latency on the first cache miss matters; the
project needs guaranteed offline operation; package registries are
themselves a privacy or supply-chain risk.

\textbf{Cost profile.} Very low upfront cost, low maintenance. Latency
spikes on cache miss; otherwise local-fast. Token cost low because only
needed files are loaded.

\subsection{Strategy 3 --- Repo packing (snapshot the project into a
single artifact)}\label{sec-strat-pack}

\textbf{Mechanism.} A tool (Repomix, code2prompt, gitingest,
files-to-prompt, yek, ai-digest, gpt-repository-loader) walks the
repository, applies ignore rules, and produces one large markdown or
XML blob that the user pastes (or the agent loads) as a single prompt
attachment.

\textbf{Wins when:} the agent has no filesystem or shell tools (web
chat interfaces, hosted assistants without code access); the task is a
one-shot Q\&A about a repo (``what does this project do?''); the team
wants a reproducible artifact tied to a specific commit.

\textbf{Loses when:} the repo is larger than the pack will compress
into; the model has limited effective context length; the task requires
multi-turn agent loops (the packed snapshot goes stale immediately on
any edit); position-sensitive attention biases hurt the buried files.

\textbf{Cost profile.} Cheap per snapshot, expensive per token in
context. Zero determinism over time (one snapshot is deterministic; a
re-pack on a later commit is not).

\subsection{Strategy 4 --- Instruct (prose rules and skills)}
\label{sec-strat-instruct}

\textbf{Mechanism.} Project-level instruction files
(\texttt{AGENTS.md}, \texttt{CLAUDE.md}, \texttt{GEMINI.md},
\texttt{.cursor/rules}, Anthropic Agent Skills) tell the agent how this
codebase wants to be worked on. Skills are progressive-disclosure
units: a short \texttt{SKILL.md} that the agent reads first, with
deeper procedural files the agent loads only when needed.

\textbf{Wins when:} the knowledge is procedural (``always run the
linter before committing''), conventional (``we use snake\_case in
Python, camelCase in TypeScript''), or stable (``never modify
\texttt{vendor/}''); the team wants a single canonical place for
``things every agent should know.''

\textbf{Loses when:} the rules are wording-sensitive (Vercel's Next.js
evals: skill-based docs lookup only worked with explicit instruction;
default skill triggering was ignored); the rules drift away from the
code as the project evolves; the model decides not to follow them.

\textbf{Cost profile.} Almost free to author and maintain. Always-on
prompt overhead is small if rules are concise; large if rules are bloated.

\subsection{Strategy 5 --- Pre-built index (semantic, symbolic, or
graph)}\label{sec-strat-index}

\textbf{Mechanism.} A background process indexes the codebase
(embeddings, BM25, AST, LSP symbols, code graph) and exposes a search
interface to the agent. The agent searches the index instead of reading
files directly. Implementations: Cursor's built-in index, Sourcegraph's
code graph (SCIP), Augment's enterprise index, Cody, Sourcebot, Bloop,
Tabnine, Glean, Continue.

\textbf{Wins when:} the codebase is large enough that grep is slow or
the search query is semantic rather than lexical; cross-file
relationships matter; the same agent will run many tasks against the
same codebase and the index is amortized.

\textbf{Loses when:} the codebase changes rapidly and the index drifts;
the index is hosted in the cloud and the codebase is sensitive; setup
cost is large for one-shot use.

\textbf{Cost profile.} High upfront index build and ongoing
re-indexing. Low per-request latency once warm. Token cost low (only
hits are loaded).

\subsection{Strategy 6 --- Annotate (overlays from external
systems)}\label{sec-strat-annotate}

\textbf{Mechanism.} Context that does not live in the repo --- code
review threads, Jira issues, design docs, Slack discussions, customer
feedback --- is exposed as overlays linked to lines, symbols, or files.
OpenCtx (now relatively dormant), NLWeb, and Sourcegraph's code-graph
context retrieval are the canonical patterns.

\textbf{Wins when:} the missing context is institutional knowledge that
explains \emph{why} code looks the way it does; the team has discipline
about putting decisions in linked systems.

\textbf{Loses when:} the overlay systems are stale; ACLs prevent the
agent from reading them; the institutional context never made it into a
queryable system.

\textbf{Cost profile.} Low per-request once integrated; integration is
itself the cost.

\subsection{Strategy 7 --- Tool-mediated agentic search}\label{sec-strat-agentic}

\textbf{Mechanism.} The agent has bash, grep, file reads, edit, and
sometimes web tools, and the model decides what to inspect each turn.
There is no pre-built index --- the model navigates the filesystem in
real time, often with reasoning between each tool call. Claude Code,
Cline, OpenHands, SWE-agent, aider, and many MCP-enabled CLIs work this
way.

\textbf{Wins when:} local files are available; the codebase changes
faster than an index can re-build; the model is strong enough to plan
multi-step searches; freshness matters (uncommitted changes, recent
edits).

\textbf{Loses when:} the search loop drags on (latency and token cost
blow up); the task requires cross-file structural reasoning that grep
cannot produce; the model spirals through file reads without
converging.

\textbf{Cost profile.} Zero upfront cost (you already have a shell).
High latency and high token cost per task. Lowest determinism: two
runs can search different files in different orders.

\subsection{Strategies compose}\label{sec-strat-compose}

In real deployments the seven rarely run alone. A typical Claude Code
setup combines instruct (CLAUDE.md), tool-mediated (bash + grep +
read), skill (progressive disclosure), and lazy fetch via MCP. Cursor
combines index (default), retrieve (\texttt{@docs}), and instruct
(\texttt{.cursor/rules}). Effect's git-subtree recommendation is pure
vendor leveraging the agent's tool-mediated capability over the
embedded source.

Figure~\ref{fig-latency-profile} compares the seven strategies on four
cost axes (per-request latency, token cost, upfront cost,
non-determinism), so a strategy can be picked by which costs the task
can afford.

\section{The Tools}\label{sec-tools}

What follows is a May 2026 snapshot of the products and projects that
implement each strategy. Tool counts and product features age fast;
treat this as a navigation aid, not a static reference. The full
extended snapshot lives in Appendix~\ref{sec-tools-appendix}.

\subsection{IDE-native context engines}\label{sec-tools-ide}

These are the products developers spend most hours inside. Each one
bakes in a default combination of strategies.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.22}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.30}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.48}}@{}}
\caption{IDE-native context engines and the strategies they
implement.}\label{tbl-tools-ide}\tabularnewline
\toprule\noalign{}
\textbf{Tool} & \textbf{Default strategy mix} & \textbf{Where it fits best}\\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\textbf{Tool} & \textbf{Default strategy mix} & \textbf{Where it fits best}\\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Cursor & Index + instruct (\texttt{.cursor/rules}) + retrieve (\texttt{@docs}) & Day-to-day work in a large, stable codebase that benefits from amortized indexing\\
Claude Code & Tool-mediated agentic + instruct (CLAUDE.md) + skills + lazy MCP fetch & High-context single-task work where freshness and exploration matter more than pre-built indices\\
GitHub Copilot & Index + tool-mediated (Copilot Chat agent mode) & Teams already deep in the GitHub ecosystem; PR-and-IDE workflows\\
Augment Code & Enterprise index + tool-mediated; 100M+ LOC repos; on-prem option & Highly regulated or massive proprietary codebases\\
Sourcegraph Amp & Code graph (SCIP) + tool-mediated; multi-repo & Cross-repo enterprise navigation\\
JetBrains AI / Junie & Index + tool-mediated, IDE-native & Teams committed to JetBrains IDEs\\
Cline / Roo Code / Kilo Code & Tool-mediated agentic via MCP & VS Code users who want bring-your-own-model\\
aider & Tool-mediated agentic + tree-sitter repo map + git-native commits & Terminal-first developers who want repo-aware AI pair programming\\
Continue & Open-source IDE assistant; pluggable & Organizations that need fully auditable open-source tooling\\
Devin / Factory Droid / OpenHands & Autonomous agent with managed env & Long-horizon autonomous tasks where the agent runs unattended\\
Warp (Agent Mode) & Tool-mediated + WarpGrep RL retriever & Terminal-centric workflows; large local repos\\
Amazon Q Developer / Kiro & Index + instruct, on AWS Bedrock & Teams already standardized on AWS\\
\end{longtable}

\subsection{MCP servers and standalone fetchers}\label{sec-tools-mcp}

MCP servers expose context as tools the agent can call. The 2026 MCP
ecosystem includes 10{,}000+ public servers; the table covers the
ones most relevant to coding agents.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.22}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.30}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.48}}@{}}
\caption{MCP servers and standalone context fetchers.}\label{tbl-tools-mcp}\tabularnewline
\toprule\noalign{}
\textbf{Server / tool} & \textbf{Strategy implemented} & \textbf{Where it fits best}\\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\textbf{Server / tool} & \textbf{Strategy implemented} & \textbf{Where it fits best}\\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Context7 & Pre-built per-library docs index over 33K+ libraries & Public-API \& migration tasks; version-matched docs\\
filesystem MCP & Tool-mediated access to a sandboxed local FS & Universal building block for agentic search\\
GitHub MCP / GitLab MCP & Tool-mediated access to issues, PRs, code & Tasks that need repo metadata beyond the working tree\\
DeepWiki MCP & Pre-built docs/wiki index from public repos & ``What does library X do?'' research\\
mgrep / Serena MCP & Symbol/AST/LSP retrieval (Strategy 5, structural variant) & Cross-file structural questions; languages with LSP support\\
Mem0 / Letta / Mnemos / Pieces LTM-2 & Long-term memory store (Strategy 6, annotate-style) & Multi-session work that needs continuity\\
OpenSrc & Lazy package fetcher into \texttt{\~{}/.opensrc} & On-demand source inspection without vendoring\\
Exa MCP / Tavily MCP / Perplexity MCP & Web search exposed as a tool & Tasks that need fresh web evidence beyond the repo\\
\end{longtable}

\subsection{Embeddings, RAG, and code search components}\label{sec-tools-embed}

These are substrates inside other tools more often than end-user
products. They get picked when teams build their own retrieval stack.

\begin{itemize}
\tightlist
\item
  \textbf{Code-trained embeddings:} voyage-code-3, Jina code embeddings,
  mxbai-embed-large, CodeBERT, GraphCodeBERT, UniXcoder, CodeT5+,
  CodeSage. Pick by language coverage and corpus size; benchmark on
  your own data because cross-domain transfer is uneven.
\item
  \textbf{Vector databases:} ChromaDB, Qdrant, Pinecone, Weaviate,
  Milvus, pgvector. The pick rarely matters at small scale; at large
  scale, focus on incremental updates and ACL support.
\item
  \textbf{Code search engines:} Sourcebot, Zoekt, Livegrep, Hound,
  ast-grep, Comby, Bloop, Greptile, Tabby. Pick AST/structural over flat
  embedding wherever language tooling permits (Pillar C evidence in
  \S\ref{sec-evidence}).
\item
  \textbf{Indexing pipelines:} CocoIndex, Indexify, FreshStack-style
  fresh document retrieval. Pick when freshness or incremental updates
  are first-class.
\end{itemize}

\subsection{Repo packaging and snapshots}\label{sec-tools-pack}

\textbf{Repomix, code2prompt, gitingest, files-to-prompt, yek,
ai-digest, gpt-repository-loader.} Each walks the repo, applies ignore
rules, and emits a single artifact. Pick by language support, default
ignore behavior, and whether you need XML, markdown, or JSON output.
None of these is wrong; they differ in defaults, not in capability.

\subsection{Multi-repo and org context}\label{sec-tools-multi}

\textbf{RepoSwarm, Backstage TechDocs, Roadie RAG, Glean, Sourcegraph,
DeepWiki, Unblocked.} The defining constraint here is ACLs and
multi-repo scale --- ordinary IDE engines do not handle either well.
Pick when ``the right context is in another repo'' is a routine
problem.

\subsection{Vendoring wrappers}\label{sec-tools-vendor}

\textbf{ingraft, Vendir (Carvel), git subtree directly, git submodule
directly, git clone with ignored directory.} These differ in how they
handle updates, ignore-pattern injection, and provenance, not in what
they store. \S\ref{sec-ingraft} covers ingraft's specific slot.

\subsection{Adjacent safety, review, and workflow tools}\label{sec-tools-adjacent}

\textbf{Qodo Merge, CodeRabbit, Greptile review mode, Sourcery review,
GitGuardian ggshield AI Hook, MCP-omnisearch, GPTScript, Jujutsu,
agentjj.} These are not primary context providers but they consume,
filter, or audit the context routed to agents during PR review,
secret detection, search aggregation, or VCS workflows.

\section{The Fit Guide}\label{sec-fit}

This is the chapter the rest of the guide exists for: given a concrete
task and a set of constraints, which combination of strategies and
tools should the reader reach for? The decision matrix
(Table~\ref{tbl-fit-matrix}) maps task class to recommended primary and
supporting strategies, and the worked scenarios that follow show each
combination playing out in practice.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.22}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.28}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.25}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.25}}@{}}
\caption{Decision matrix --- task class to recommended strategy
combination.}\label{tbl-fit-matrix}\tabularnewline
\toprule\noalign{}
\textbf{Task class} & \textbf{Primary strategy} & \textbf{Supporting strategies} & \textbf{Representative tools}\\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\textbf{Task class} & \textbf{Primary strategy} & \textbf{Supporting strategies} & \textbf{Representative tools}\\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Learn an idiom-heavy framework & Vendor & Tool-mediated agentic + instruct & ingraft / git subtree + Claude Code\\
Major version migration & Docs index (instruct) & Lazy fetch + tool-mediated & Context7 / AGENTS.md docs index + Claude Code or Cursor\\
Onboard to large unfamiliar codebase & Pre-built index & Tool-mediated + repo map & Cursor or Sourcegraph Amp + aider repo-map\\
Cross-repo enterprise refactor & Multi-repo index (Strategy 5+6) & Tool-mediated & Sourcegraph Amp / Glean / Augment\\
Debug a flaky / failing test & Tool-mediated agentic & Lazy fetch for package source & Claude Code / Cline / aider + OpenSrc\\
One-shot Q\&A on an unfamiliar repo & Repo pack & --- & Repomix / gitingest\\
Security review on a sensitive codebase & On-prem index + instruct & Tool-mediated, agentic search disabled & Augment / Tabnine on-prem; strict trust policy\\
Test repair after a refactor & Tool-mediated agentic & Symbol/AST index & Claude Code + Serena MCP / mgrep\\
\end{longtable}

\subsection{Scenario 1 --- Learn an idiom-heavy
framework}\label{sec-fit-scenario-effect}

\textbf{Situation.} The team starts a TypeScript service that depends
heavily on Effect, a framework whose human-facing docs are precise but
deliberately abstract; the patterns live in the source.

\textbf{Recommended bundle.} Vendor (Strategy~1) the Effect source into
\texttt{vendor/effect}, mark it read-only via \texttt{AGENTS.md}
(Strategy~4), and use Claude Code or Cursor (Strategy~7,
tool-mediated) over the embedded source. The agent finds idioms by
grep + read on real code and tests. Pulled docs (Strategy~2 lazy
fetch) cover migration questions when the source view is not enough.

\textbf{Why this wins.} Source-over-docs evidence (Brown 2026) is
strongest for idiom-heavy frameworks. The vendored copy gives the
agent navigable, version-matched patterns with zero per-request
latency or token cost. The \texttt{AGENTS.md} rule keeps humans from
editing vendored files by mistake.

\textbf{Antipatterns.} Pasting Effect docs into the prompt every turn;
indexing a third-party library that changes weekly; relying on the
model's pre-training memory of Effect 2.x while the project uses 3.x.

\subsection{Scenario 2 --- Major version migration (Next.js
15 \texorpdfstring{$\to$}{->} 16)}\label{sec-fit-scenario-nextjs}

\textbf{Situation.} A team must upgrade a Next.js 15 app to Next.js
16. Hundreds of small API changes, several breaking ones, and the
existing code uses deprecated hooks.

\textbf{Recommended bundle.} Always-on docs index (Strategy~4, the
Vercel pattern): a compressed \texttt{AGENTS.md} that points at a
local \texttt{.next-docs/} directory with version-matched migration
docs. Pair with Strategy~5 (symbol/AST index of the local codebase) so
the agent can locate every deprecated API call by symbol, not by
substring. Tool-mediated agentic (Strategy~7) does the edits.

\textbf{Why this wins.} Vercel's own Next.js eval suite
(Figure~\ref{fig-vercel-eval}): the \texttt{AGENTS.md} docs-index
pattern scored 100\% on their hardened eval, versus 53\% baseline and
79\% with skill plus explicit instruction (Gao 2026a). Migration tasks
are public-API tasks; docs are authority, source is not.

\figureVercelEval

\textbf{Antipatterns.} Vendoring Next.js source (huge, generated,
distractor-heavy); web-searching every API change (stale, no version
guarantee); relying on the model's pre-training memory of an older
Next.js.

\subsection{Scenario 3 --- Onboarding to a large unfamiliar
codebase}\label{sec-fit-scenario-onboard}

\textbf{Situation.} A new engineer joins a 2M-LOC monorepo.

\textbf{Recommended bundle.} Pre-built index (Strategy~5, semantic +
structural) for orientation; tool-mediated agentic (Strategy~7) for
deep dives; aider-style tree-sitter repo map (Strategy~3,
pack-compressed) for ``which files matter for X''. Instruct
(Strategy~4) project rules via \texttt{AGENTS.md} or
\texttt{CLAUDE.md} so the agent inherits team conventions.

\textbf{Why this wins.} 2M LOC is past the break-even point where
index maintenance pays off: indexing wins over agentic search when
repository size, query frequency, and cross-file complexity are large
enough to amortize the build and re-indexing cost
($C_\mathrm{toolCalls} + C_\mathrm{tokens} + C_\mathrm{latency} >
C_\mathrm{indexBuild} + C_\mathrm{indexDrift} +
C_\mathrm{retrievalErrors}$). Repo maps are the right orientation tool
because exact code is too long for orientation and architecture
diagrams drift.

\textbf{Antipatterns.} Repo-packing 2M LOC into a single prompt;
forcing agentic search to grep across the entire monorepo on every
turn; skipping \texttt{AGENTS.md} and re-learning conventions task by
task.

\subsection{Scenario 4 --- Cross-repo enterprise
refactor}\label{sec-fit-scenario-multi}

\textbf{Situation.} A platform team must rename a deprecated API used
across 40 internal repos.

\textbf{Recommended bundle.} Multi-repo managed index (Strategy~5 +
Strategy~6 annotate, Glean / Sourcegraph / Augment) with ACL-aware
search. Tool-mediated agentic (Strategy~7) per repo to do the edits.
Instruct (Strategy~4) the policy via \texttt{AGENTS.md} in each repo
so the agent picks up project-specific overrides.

\textbf{Why this wins.} Org context requires ACLs, multi-repo scale,
and governance that IDE engines do not handle well.

\textbf{Antipatterns.} Doing it from a single IDE with no cross-repo
visibility; using a public cloud index for proprietary code.

\subsection{Scenario 5 --- Debug a flaky / failing
test}\label{sec-fit-scenario-debug}

\textbf{Situation.} Test \texttt{TestPaymentRetry} fails intermittently
on CI.

\textbf{Recommended bundle.} Tool-mediated agentic (Strategy~7) is the
primary: the agent reads logs, greps the test, follows call chains,
inspects the package source if needed. Lazy fetch (Strategy~2)
provides on-demand package source for any third-party retry library.
No pre-built index needed.

\textbf{Why this wins.} Debugging is exploratory by nature; pre-built
indexes go stale faster than uncommitted changes; the cost of agentic
search is amortized over a short interactive session, not 40
production tasks.

\textbf{Antipatterns.} Indexing for a one-off debug; vendoring a
dependency just to read three functions; repo-packing the entire
codebase when the bug is in two files.

\subsection{Scenario 6 --- One-shot Q\&A on an unfamiliar
repo}\label{sec-fit-scenario-oneshot}

\textbf{Situation.} An engineer wants to ask ``what does this
open-source repo do, and how does its auth module work?'' from a web
chat interface with no shell tools.

\textbf{Recommended bundle.} Repo pack (Strategy~3) with Repomix or
gitingest, paste into the chat.

\textbf{Why this wins.} Repo packs are the right tool when (a) the
agent has no filesystem and (b) the task is one-shot. The downsides
(staleness, position bias) do not matter for a single Q\&A.

\textbf{Antipatterns.} Setting up an MCP server, an index, and a
vendoring pipeline for a question that takes 60 seconds.

\subsection{Scenario 7 --- Security-sensitive
codebase}\label{sec-fit-scenario-security}

\textbf{Situation.} A regulated team works on financial-services code
where source must not leave the corporate network.

\textbf{Recommended bundle.} On-prem index (Strategy~5; Augment VPC /
Tabnine air-gapped / self-hosted Sourcegraph) + instruct (Strategy~4)
with strict trust policy that disables web fetch and external MCP
servers. Tool-mediated agentic (Strategy~7) only against local files.

\textbf{Why this wins.} Privacy constraints rule out cloud retrieval.
The on-prem index keeps the productivity gain without the data egress.
Strict trust policy on \texttt{AGENTS.md} closes the prompt-injection
surface.

\textbf{Antipatterns.} Sending repo content through any third-party
LLM that retains training data; using public MCP servers without
auditing them; relying on developer discipline alone to keep secrets
out of prompts.

\section{The Evidence --- Why These Choices Matter}\label{sec-evidence}

The recommendations in \S\ref{sec-strategies}--\S\ref{sec-fit} are
shaped by a decade of research on long-context models, retrieval, and
agent scaffolding. This chapter is the supporting layer: the studies
and benchmarks that explain \emph{why} the strategies break the way
they do. A reader who trusts the recommendations can skim it; a reader
who wants to argue with them should not.

The original section title was ``Why Context Allocation Is Hard,'' and
that framing still organizes the material into four pillars: long
context degrades, retrieval is conditional, structure beats flat, and
distractors actively hurt.

\subsection{Context windows are not attention
guarantees}\label{sec-attention}

Long windows increase storage, not reliable computation over every
token. N. F. Liu et al. (2024) established the now-canonical positional
failure: models perform best when relevant information is near the
beginning or end of the context and degrade when evidence is buried in
the middle. {Hsieh et al.} (2024) showed that advertised context length
can overstate effective context length, especially on multi-hop
synthetic tasks.

LongCodeBench extends the problem to realistic code comprehension and
repair (Rando et al. 2025). It evaluates LongCodeQA and LongSWE-Bench
across 32K, 64K, 128K, 256K, 512K, and 1M context brackets where
supported. The two tasks have to be reported separately --- collapsing
them into one ``LongCodeBench accuracy'' number is the most common
misreading of the paper, and it obscures the result that matters.

\begin{quote}
On \textbf{LongSWE-Bench} (executable bug-fix pass rate on real GitHub
issues), Claude 3.5 Sonnet solved 29\% of issues at 32K and 3\% at 256K;
it was not evaluated at 512K or 1M. Gemini 2.5 Pro held 22--25\% through
256K and only collapsed at 512K/1M (12\%, 7\%). GPT-4.1 sat near 1\% at
every bracket. Qwen2.5, Llama 3.1, and Llama 4 Scout each solved zero
issues at every bracket. On \textbf{LongCodeQA} (multiple-choice
comprehension), Qwen2.5 peaked at 70.2\% at 512K and fell to 40.0\% at
1M, while GPT-4.1 and Gemini 2.5 Pro showed no comparable cliff.
\end{quote}

\figureLongContextDegradation

Three observations follow that single context degradation is not enough
to characterize. First, the cliff is \emph{not} monotonic across models:
Gemini 2.5 Pro on LongSWE-Bench is the clean counterexample, and GPT-4.1
on LongCodeQA actively improves with more context. Second, the cliff is
\emph{not} the same across tasks: the ranking that picks the best
1M-token comprehender and the ranking that picks the best 256K-token
bug-fixer are different rankings. Third, the often-quoted ``Claude
collapses from 29\% to 3\% at 1M'' headline silently extrapolates beyond
the published 256K bracket; the honest version is 32K to 256K. A router
that picks a single ``effective length'' for all models and all tasks is
discarding the part of the result that has operational consequences.

Figure~\ref{fig-longcodebench-heatmap} extends the same data to the full
grid of ten models across six brackets and two tasks. The bug-fix grid is
mostly red --- four models score zero at every bracket; the comprehension
grid is mostly green for GPT-4.1, Gemini 2.5 Pro, and Llama 4 Scout. The
visual contrast is the operational point: the model that comprehends well
at 1M and the model that repairs code well at 256K are not the same model.

\figureLongCodeBenchHeatmap

The newer context-rot literature sharpens the same point. Chroma's
technical report evaluates 18 models and isolates cases where increasing
input length alone makes outputs less reliable, including semantic
needle variants, distractors, LongMemEval comparisons, and repeated-word
copying tasks (Hong, Troynikov, and Huber 2025). NoLiMa extends
needle-in-a-haystack beyond literal matching and reports severe drops
when questions and needles have little lexical overlap, even when short
contexts are nearly solved (Modarressi et al. 2025). AbsenceBench adds a
different failure class: models that can retrieve present evidence may
still fail to identify missing information, including omitted lines in
GitHub pull-request diffs (Fu et al. 2025).

The defensible conclusion is not ``1M always fails.'' The conclusion is
that advertised context length is an upper bound, not an operating
target. Routing budgets should be calibrated per model, task class, and
representation, ideally by local evals or conservative defaults.

\subsection{Retrieval can help or hurt}\label{sec-retrieval-hard}

{Z. Z. Wang et al.} (2025) asks directly whether retrieval can augment
code generation. Its answer is conditional: high-quality retrieved
contexts improve some settings, especially weaker models and basic
programming tasks, while retrievers still struggle when lexical overlap
is limited and generators do not always integrate extra context
effectively. Asai et al. (2024) reaches a similar architectural lesson
from a broader RAG setting: indiscriminate fixed retrieval can reduce
versatility and produce unhelpful generations; retrieval should be
adaptive.

Figure~\ref{fig-retrieval-conditional} crystallizes the empirical pattern.
The same retrieval mechanism is a clear win in one quadrant
(weaker models on basic tasks --- CodeRAG-Bench reports +15.6 to +17.8 pp
for StarCoder2-7B on HumanEval and MBPP) and a clear loss in the opposite
quadrant (frontier reasoning models hit by semantic distractors). The
operational consequence is that ``do we retrieve?'' has to be evaluated
against \emph{the agent's model strength and the task's complexity
together,} not as a global default.

\figureRetrievalConditional

The router should not ask ``RAG or not?'' It should ask which retrieval
substrate has the lowest expected miss, distractor, staleness, and
latency cost for this task.

That means semantic and non-semantic search should be treated as
different instruments, not rival slogans. BM25 remains a strong lexical
baseline because exact terms, identifiers, and error strings often are
the task (Robertson and Zaragoza 2009). Learned sparse retrieval such as
SPLADE keeps lexical auditability while adding expansion terms (Formal,
Piwowarski, and Clinchant 2021). Dense retrieval and text embeddings are
useful when intent does not share surface form with the target, but
benchmarks such as BEIR and MTEB show that retrieval quality is highly
task- and distribution-dependent (Thakur et al. 2021; Muennighoff et al.
2023). CodeSearchNet made the natural-language-to-code gap explicit for
semantic code search (Husain et al. 2019), while GraphCodeBERT shows why
code structure and data flow matter beyond flat token embeddings (Guo et
al. 2021). Late-interaction retrievers such as ColBERT occupy a useful
middle ground: they retain token-level matching signals while allowing
document representations to be precomputed (Khattab and Zaharia 2020).
ColBERTv2 and PLAID make the deployment point stronger by reducing
late-interaction storage and serving cost without giving up the core
matching advantage (Santhanam et al. 2022a; Santhanam et al. 2022b).
Mixedbread's mxbai work is relevant as practitioner evidence for both
dense embeddings and ColBERT-style reranking, and mgrep is a concrete
coding-agent product claim that semantic search can reduce token usage
relative to repeated grep loops while still preserving grep as the exact
match fallback (Lee et al. 2024a; Lee et al. 2024b; Mixedbread 2026).
Multi-function embedding systems such as BGE-M3 push in the same
direction by supporting dense, sparse, and multi-vector retrieval modes
inside one model family (Chen et al. 2024).

Recent code and technical-document retrieval benchmarks make the same
lesson more concrete for software agents. FreshStack argues for
benchmarks built from recent technical documentation, where off-the-shelf
retrievers can lag far behind oracle context on niche stacks (Thakur et
al. 2025). COIR separates code retrieval into multiple task families
rather than treating ``code search'' as one problem (Li et al. 2024).
CoSIL frames issue localization as graph-guided repository search rather
than flat chunk retrieval (Jiang et al. 2025). For routing, these papers
matter because they move the selector question from ``semantic search or
grep?'' to ``which retrieval substrate matches this task, authority, and
failure mode?''

\figureRetrievalLandscape

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.5000}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.5000}}@{}}
\caption{Retrieval substrates and their common failure
modes.}\label{tbl-retrieval-types}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Retrieval type
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Typical failure mode
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Retrieval type
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Typical failure mode
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Lexical grep & Misses semantic matches and renamed concepts \\
Semantic vector search & Retrieves plausible but wrong chunks \\
Graph retrieval & Depends on graph quality and language support \\
AST chunking & Can miss runtime behavior and dynamic dispatch \\
Docs retrieval & Docs may be stale, incomplete, or for the wrong
version \\
Agentic search & Slow, non-deterministic, and tool-call heavy \\
Hybrid search & More accurate but harder to debug and reproduce \\
\end{longtable}

\subsection{Structure matters}\label{sec-structure}

Repository-level work consistently shows that flat text is a weak
abstraction for code. F. Zhang et al. (2023), Shrivastava et al. (2023),
Ding et al. (2024), {Ding et al.} (2023), {Liang et al.} (2024), {Phan
et al.} (2025), and {W. Liu et al.} (2024) all point toward the same
lesson: imports, definitions, call relationships, type information,
symbols, and file structure provide routing signals that raw chunks do
not. {Y. (Sherry). Zhang et al.} (2025) adds a newer preprint result for
AST-aware chunking.

The careful implication is not ``AST plus graph must always be the
default.'' It is: for repository-scale retrieval, the prior should favor
structural retrieval whenever the language ecosystem and latency budget
permit it.

Figure~\ref{fig-structural-gains} collects nine reported gains from six
independent studies. The metric varies across studies (Recall@5, Pass@1,
exact match, top-1 precision, identifier match), so the bars are not
strictly comparable, but every published comparison points the same
direction. Even the smallest cAST result (+2.7 percentage points on
SWE-bench Pass@1) materially shifts real bug-fix rates; the larger
relative gains (CoCoMIC +33.94\%, RepoHyper +49\%, REPOFUSE +50\%,
Code-Craft +82\%) are large enough that the question for the router stops
being ``does structure help?'' and becomes ``which structural substrate
fits this language and latency budget?''

\figureStructuralRetrievalGains

\subsection{Wrong context has concrete harms}\label{sec-wrong-context}

Context quality fails through recognizable mechanisms.
Table~\ref{tbl-failure-modes} maps each failure mode to a router
response.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}@{}}
\caption{Failure modes that motivate
routing.}\label{tbl-failure-modes}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Failure mode
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Description
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Router response
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Failure mode
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Description
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Router response
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Attention dilution & Relevant tokens are present but weakly attended &
Reduce bundle size; place high-value context early or late \\
Distractor interference & Plausible irrelevant context changes the
reasoning path & Prune aggressively; prefer version-specific
authority \\
Representation mismatch & Docs explain concepts but not usage patterns &
Fetch examples, source, or tests \\
Version skew & Retrieved docs/source do not match installed package &
Resolve through lockfile and provenance \\
Retrieval miss & Search fails to fetch the necessary file or symbol &
Expand via graph, AST, LSP, or agentic search \\
Context staleness & Index reflects old repo state & Prefer live grep or
incremental index \\
Trust contamination & External docs contain malicious or irrelevant
instructions & Quote, sanitize, label provenance, sandbox tools \\
Cost blowup & Agent spends many turns searching & Precompute repo maps
or local indexes \\
\end{longtable}

\figureFailureModes

Examples are easy to find in everyday agent work: the agent uses latest
docs for an older installed package, imports a private symbol after
reading source, acts on a StackOverflow answer for the wrong major
version, searches a stale index while local uncommitted changes matter,
follows instructions embedded in third-party docs, or receives a whole
repo pack and overlooks the relevant file in the middle. These are not
model failures alone. They are allocation failures.

\subsection{Context changes during the agent loop}\label{sec-phases}

Context routing is online, not a one-time preprocessing step. Each phase
of an agent loop wants different material.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.5000}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.5000}}@{}}
\caption{Task-phase routing.}\label{tbl-phases}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Phase
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Useful context
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Phase
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Useful context
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Orient & AGENTS.md, repo map, architecture summary, task brief \\
Locate & Code search, symbols, imports, stack traces, issue
references \\
Plan & Relevant local files, docs, examples, package conventions \\
Edit & Exact files, types, tests, local wrappers, public API
contracts \\
Verify & Test logs, CI output, type errors, runtime traces \\
Repair & Failing trace, package source, issue history, migration docs \\
Summarize & Decision log, changed files, unresolved risks, provenance \\
\end{longtable}

\figurePhaseHeatmap

This aligns with Anthropic's context-engineering framing: curation and
maintenance of useful tokens happens throughout inference, not only at
prompt construction time (Anthropic 2025a). Emerging OSS-focused context
engineering work makes the same point at repository scale: agent
configuration files, guidance documents, and tool manifests vary widely,
which means context policy itself becomes part of the software
architecture rather than an afterthought (Mohsenimofidi et al. 2025).

\section{How to Think About Routing}\label{sec-routing}

When the decision matrix in \S\ref{sec-fit} does not cover the case, the
reader needs a framework for reasoning from scratch. This chapter is
that framework: five orthogonal axes that any routing decision can be
described along, followed by the mechanics of an explicit routing
policy (inputs, scoring, authority, provenance, trust, explainability).

The axes are the same five used by every coding-agent stack on the
market, whether or not their docs name them: where context lives
(\emph{placement}), what form it takes (\emph{representation}), what
chooses it (\emph{selector}), when it loads (\emph{timing}), and which
constraint binds (\emph{dominant constraint}). Naming them makes a
fuzzy decision into a sequence of small, defensible decisions.

Before the axes, one principle holds across all of them: the operational
goal is \textbf{smallest sufficient context}, not maximum context.
Sufficiency means the agent can identify the relevant authority,
produce a correct change, and verify it without materially misleading
distractors. Smallest does not mean fewest tokens; it means lowest
total cost across tokens, latency, privacy, staleness, trust, and
distractor risk. Figure~\ref{fig-utility-tradeoff} shows the typical
curve: utility climbs as relevant context is added, plateaus, then
falls as the bundle grows past the model's effective attention.

\figureUtilityTradeoff

\subsection*{The five axes}
\addcontentsline{toc}{subsection}{The five axes}

The earlier ``seven primitives'' vocabulary is useful as shorthand, but
it mixes dimensions. Vendoring, indexing, retrieving, packing, skills,
annotations, and agentic reads are common bundles across several
independent axes. A more useful taxonomy separates the axes first.

\subsection{Axis 1: placement}\label{sec-axis-placement}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.5000}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.5000}}@{}}
\caption{Where context lives.}\label{tbl-placement}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Placement
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Examples
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Placement
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Examples
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
In prompt & Pasted file, repo pack, selected snippets \\
On disk in workspace & Vendored source, cloned dependency, exported
docs \\
Local cache & OpenSrc-style package cache, tool-managed snapshots \\
Local index & Sourcebot, Serena, local vector DB, SCIP/LSP graph \\
Remote service & Context7, Augment, Glean, hosted semantic search \\
Model or tool memory & Skills, summaries, persistent notes, agent
memory \\
\end{longtable}

\subsection{Axis 2: representation}\label{sec-axis-representation}

\begin{longtable}[]{@{}ll@{}}
\caption{How context is
represented.}\label{tbl-representation}\tabularnewline
\toprule\noalign{}
Representation & Examples \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
Representation & Examples \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Raw source & Vendored repository, fetched package source \\
Human docs & API docs, README, migration guides, changelogs \\
Symbol graph & SCIP, LSP, AST, call/reference graph \\
Embedding chunks & Vector search over files, docs, or issues \\
Summaries & Repo maps, architecture notes, skills, memory \\
Executable tools & MCP tools, scripts, local CLI commands \\
Instructions & AGENTS.md, CLAUDE.md, project rules \\
\end{longtable}

\subsection{Axis 3: selector}\label{sec-axis-selector}

\begin{longtable}[]{@{}ll@{}}
\caption{Who or what selects
context.}\label{tbl-selector}\tabularnewline
\toprule\noalign{}
Selector & Examples \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
Selector & Examples \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Human & \texttt{@file}, explicit package, manual doc link \\
Static policy & allowlist routes \texttt{effect} to vendor \\
Heuristic router & import count, package size, privacy class \\
Learned retriever & embedding search, reranker \\
Graph retriever & symbol, dependency, call/reference expansion \\
Agentic search & \texttt{grep}/read/bash loop \\
Oracle/evaluator & benchmark-only upper bound \\
\end{longtable}

\subsection{Axis 4: timing}\label{sec-axis-timing}

\begin{longtable}[]{@{}ll@{}}
\caption{When context is loaded.}\label{tbl-timing}\tabularnewline
\toprule\noalign{}
Timing & Examples \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
Timing & Examples \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Always loaded & AGENTS.md, compact repo guidance \\
Session start & Repo map, recent decision log \\
Task start & Package docs, relevant tests, task issue \\
Just in time & Agent calls search, router fetches source \\
After failure & Fetch package source after stack trace points there \\
Compacted later & Summarize stale context for continuation \\
\end{longtable}

\subsection{Axis 5: dominant constraints}\label{sec-axis-constraints}

\begin{longtable}[]{@{}ll@{}}
\caption{Constraints that dominate route
choice.}\label{tbl-constraints}\tabularnewline
\toprule\noalign{}
Constraint & Examples \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
Constraint & Examples \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Token budget & Long-context degradation, prompt cost \\
Latency & Agentic search and remote retrieval delays \\
Privacy & Local-only code, no cloud index, no remote docs \\
Freshness & Local changes, docs drift, moving issues \\
Trust & MCP/tool/skill supply chain, prompt injection \\
Determinism & Reproducible context versus live retrieval \\
Version specificity & Lockfile-resolved package versions \\
\end{longtable}

The old primitives can now be expressed more precisely:

\begin{itemize}
\tightlist
\item
  \textbf{Vendoring} = workspace placement + raw source representation +
  static or heuristic selection + task-time loading + strong
  reproducibility + local privacy.
\item
  \textbf{Indexing} = local or remote placement + symbol/embedding
  representation

  \begin{itemize}
  \tightlist
  \item
    learned or graph selector + low-latency query + staleness
    management.
  \end{itemize}
\item
  \textbf{Skills} = memory placement + instruction/tool representation +
  metadata-based selector + progressive disclosure + high token
  efficiency.
\item
  \textbf{Repo pack} = prompt placement + raw or summarized
  representation + human or static selector + one-shot timing + weak
  freshness after creation.
\end{itemize}

For implementation, these axes become a route descriptor rather than a
prose taxonomy.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}@{}}
\caption{Route-descriptor axes for a context
object.}\label{tbl-route-descriptor}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Axis
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Question
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Example values
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Axis
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Question
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Example values
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Object kind & What is being routed? & Package source, repo subtree, doc
page, API spec, ticket, skill, live endpoint \\
Materialization & Where does it live for the task? & Committed, cached,
packed, mounted, remote-only \\
Representation & In what form is it exposed? & Raw source, structural
summary, prose doc, metadata envelope, search hit \\
Access path & How does the agent reach it? & Direct read, local grep,
local index, remote retrieval API, MCP tool \\
Freshness mode & How stable is it? & Pinned snapshot, refreshable cache,
live mutable \\
Authority/trust & How much should the agent rely on it? & Canonical,
upstream authoritative, derived, advisory, unverified \\
Privacy scope & What boundary constrains it? & Repo-safe, local-only,
restricted, remote-call-required \\
\end{longtable}

%====================================================================
%  CUT: Old "Practitioner Strategies as Operating Zones".
%  This material is superseded by \S\ref{sec-strategies} (The Strategies)
%  and \S\ref{sec-fit} (The Fit Guide). Preserved here in case a future
%  edit wants to pull specific quotes (Vercel evals, Effect origin,
%  source-vs-docs decision table, break-even formula).
%====================================================================
\iffalse
\section{Practitioner Strategies as Operating
Zones}\label{sec-practitioner}

The practitioner landscape is not a consensus. It is structured
disagreement. Each camp has a correct operating zone and a failure mode.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedright\arraybackslash}p{(\linewidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedright\arraybackslash}p{(\linewidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedright\arraybackslash}p{(\linewidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedright\arraybackslash}p{(\linewidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedright\arraybackslash}p{(\linewidth - 10\tabcolsep) * \real{0.1667}}@{}}
\caption{Practitioner camps, assumptions, failure modes, and routing
implications.}\label{tbl-camps}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Camp
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Claim
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Evidence
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Hidden assumption
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Failure mode
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Router implication
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Camp
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Claim
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Evidence
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Hidden assumption
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Failure mode
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Router implication
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Source vendoring & Agents learn patterns better from source than prose &
Effect subtree argument (Brown 2026) & Source is available, navigable,
version-matched, and not huge & Internal APIs, large repos, poor
comments, private access & Vendor high-centrality packages with
manageable source and high idiom value \\
Agentic search & Let the model inspect live files instead of maintaining
an index & Claude Code and Cline practice (S. (swyx). Wang and Trotta
2025; Baumann 2025) & Agent has strong tools and time budget & Latency,
non-determinism, repeated search loops & Prefer live search when local
files exist and freshness matters \\
Pre-indexing & Large repos need fast semantic or structural context &
Cursor and Sourcegraph/Amp claims (Cursor 2026; Sourcegraph 2025) &
Index is fresh, private enough, and amortized over many tasks &
Staleness, privacy, setup cost & Index when repo size and query
frequency justify maintenance \\
Repo maps & Summaries preserve structure under tight budgets & aider
tree-sitter repo map (Gauthier 2023) & Symbols/signatures are enough for
orientation & Missing details and runtime behavior & Use maps for
orientation, then expand to exact files \\
Skills and rules & Procedural knowledge should be progressively
disclosed & Agent Skills and CLAUDE.md practice (Anthropic 2025b;
HumanLayer (Kyle) 2025) & Knowledge is stable and reusable & Overbroad
instructions and hidden script risk & Use skills for procedures,
conventions, and reusable workflows \\
Always-on docs indexes & A compact project instruction file can reliably
point agents at versioned local docs & Vercel Next.js agent evals (Gao
2026a) & The index is small, accurate, and points to local docs the
agent can read & Context bloat; stale copied docs; agent ignores project
instruction files & Put high-value retrieval pointers in AGENTS.md, and
reserve skills for explicit workflows \\
Managed enterprise context & Org search needs ACLs, multi-repo scale,
and governance & Glean, Augment, Sourcegraph, DeepWiki-style systems &
Cloud or managed index is allowed & IP leakage, opaque ranking, vendor
lock-in & Route through managed systems only when policy permits \\
\end{longtable}

\subsection{Effect as the originating case}\label{sec-effect-origin}

Effect is the motivating case for ingraft and for this paper. Brown's
post makes a narrow but important source-vendoring claim: for an
idiom-heavy TypeScript framework, human-facing docs can be too lossy for
coding agents, while source, tests, examples, and module structure
preserve the patterns agents need (Brown 2026). The recommended route is
git subtree under a read-only reference directory, with agent
instructions and editor/tool ignores so humans are not overwhelmed by
vendored code.

The public article should not be treated as a controlled benchmark
report. It does not publish task definitions, model list, run counts, or
pass/fail tables. Its evidential role here is practitioner operating
experience: for Effect, the authors report that locally available source
is more useful to agents than fragmented web fetches, human docs, or
compiled \texttt{node\_modules} surfaces. That is enough to justify
source vendoring as a candidate route, not enough to make it the global
default.

This is exactly the router opening. If source vendoring wins for Effect,
the hidden variables are high package centrality, idiom density,
navigable TypeScript source, useful tests and examples, manageable
repository size, and need for version-specific source. The router
implication is not ``always vendor''; it is ``vendor when source
authority and idiom value outweigh repository impact and distractor
risk.''

\subsection{Always-on indexes versus invoked
skills}\label{sec-agents-md-evals}

Vercel's Next.js team published a useful negative result for the router
thesis: on a hardened Next.js 16 eval suite, an on-demand docs skill was
often not invoked and did not improve over the no-docs baseline, while
explicit instructions improved results but were wording-sensitive (Gao
2026a). Their best condition was a compressed \texttt{AGENTS.md} docs
index: the project file kept a small map of version-matched docs always
visible, and the agent read the actual docs from a local
\texttt{.next-docs/} directory when needed. The reported pass rates were
53\% for baseline, 53\% for default skill use, 79\% for skill plus
explicit instruction, and 100\% for the \texttt{AGENTS.md} docs index.

\figureVercelEval

This is vendor evidence, not an independent controlled study, and it is
specific to Next.js tasks and the harnesses tested. Still, it sharpens
the routing model: there are cases where the best route is not to load
full docs into context and not to rely on skill discovery, but to expose
a small always-on retrieval index that removes the agent's decision to
ask for help. Vercel's public \texttt{next-evals-oss} repository is also
a practical evaluation pattern: self-contained project fixtures with
\texttt{PROMPT.md}, withheld \texttt{EVAL.ts}, paired experiments with
and without \texttt{AGENTS.md}, and exported docs-impact deltas (Vercel
2026). The merged \texttt{@next/codemod\ agents-md} command shows the
implementation route: detect the framework version, fetch matching docs,
generate a compressed index, inject it between markers in
\texttt{AGENTS.md} or \texttt{CLAUDE.md}, and ignore the local docs
directory (Gao 2026b).

\subsection{Source is not always better than
docs}\label{sec-source-docs}

The Effect case is one operating zone, not a universal source-over-docs
rule. The source-vendoring argument is strongest when agents need
idioms, examples, type-level patterns, framework internals, and
implementation behavior. Docs are often the authority for public API
intent, migration guidance, deprecation warnings, and security
constraints.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.5000}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.5000}}@{}}
\caption{Source-versus-docs decision
rule.}\label{tbl-source-docs}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Task need
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Prefer
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Task need
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Prefer
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Learn public API quickly & Docs and curated examples \\
Understand behavior edge case & Source and tests \\
Follow idioms & Source examples plus local usage \\
Avoid private/internal APIs & Official docs and public type surface \\
Debug package internals & Version-matched source \\
Migrate across versions & Changelog and migration guide, then local
usage \\
Verify generated code & Tests plus source \\
Understand package philosophy & Docs and design notes \\
\end{longtable}

\subsection{Agentic search versus indexing is a break-even
problem}\label{sec-break-even}

The unresolved question is not whether indexing or agentic search is
``right.'' It is when the cost of maintaining an index exceeds the cost
of repeated live search.

Agentic search wins when:

\[
C_\mathrm{toolCalls} + C_\mathrm{tokens} + C_\mathrm{latency}
<
C_\mathrm{indexBuild} + C_\mathrm{indexDrift} + C_\mathrm{privacy}
+ C_\mathrm{retrievalErrors}
\]

Indexing wins when repository size, query frequency, and cross-file
complexity are large enough to amortize index maintenance. This depends
on churn, local change frequency, language tooling, privacy rules, and
whether cross-repo search matters. The router should make that
break-even explicit rather than choosing a side globally.
\fi
%====================================================================
%  END CUT (Practitioner Strategies)
%====================================================================

\subsection*{Routing in practice}\label{sec-router}
\addcontentsline{toc}{subsection}{Routing in practice}

The five axes describe \emph{where} a routing decision lives. The
remainder of this chapter describes \emph{how} an explicit policy
actually selects a route: what inputs it consumes, how it scores
candidates, how source authority resolves conflicts, how provenance
travels with the bundle, how trust policy gates the choice, and how
the route is explained to a human reviewer.

\subsection{Inputs and outputs}\label{sec-router-io}

A context router takes:

\begin{itemize}
\tightlist
\item
  task statement and task type;
\item
  repository state, including local changes and lockfiles;
\item
  knowledge object metadata;
\item
  model and tool capabilities;
\item
  privacy and trust policy;
\item
  history of failures or user overrides.
\end{itemize}

It outputs:

\begin{itemize}
\tightlist
\item
  selected route and alternatives considered;
\item
  context bundle or tool affordance;
\item
  provenance record;
\item
  trust class and intended use;
\item
  explanation and override path;
\item
  optional context-lock update.
\end{itemize}

\figureRouterArchitecture

\subsection{Strategy scoring}\label{sec-strategy-scoring}

The router can begin as a rule system, but the policy should expose
scoring features so users can understand why a route was chosen.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedright\arraybackslash}p{(\linewidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedright\arraybackslash}p{(\linewidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedright\arraybackslash}p{(\linewidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedright\arraybackslash}p{(\linewidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedright\arraybackslash}p{(\linewidth - 10\tabcolsep) * \real{0.1667}}@{}}
\caption{Scoring features across route
families.}\label{tbl-scoring}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Feature
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Vendoring
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Lazy fetch
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Docs
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Index
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Skill
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Feature
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Vendoring
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Lazy fetch
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Docs
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Index
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Skill
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Version fidelity & High if lockfile-resolved & High if resolved & Medium
to high & Varies & Low unless versioned \\
Token efficiency & Medium & Medium & High & High & High \\
Agent navigability & High & High & Medium & Medium/high & Low/medium \\
Setup cost & Medium/high & Low & Low & High & Low \\
Privacy & High & High & Varies & Varies & High \\
Freshness & Pinned & Pinned & Fresh but may mismatch & Drift risk &
Stable \\
Distractor risk & High if huge & Medium & Medium & Medium & Low \\
Best for & Deep idioms & Occasional inspection & Public API/migration &
Large repos & Conventions/procedures \\
\end{longtable}

Figure~\ref{fig-latency-profile} re-encodes the strategy choice as four
cost axes (per-request latency, token cost per task, upfront/maintenance
cost, and determinism) on a shared 1--10 scale. No strategy minimizes all
four. Strategies on the left (vendor, cache, pack) pay an upfront and
maintenance cost but recover it on latency, tokens, and determinism;
strategies on the right (MCP fetch, agentic search) are cheap to set up
and expensive per call, with agentic search the only strategy that is
expensive on \emph{both} latency and tokens while losing determinism. The
router's job is to pick the strategy whose dominant cost the task can
afford --- not the strategy that is cheapest in the abstract.

\figureLatencyProfile

Default policy rules follow from the table:

\begin{itemize}
\tightlist
\item
  Vendor when package centrality is high, source size is manageable,
  privacy matters, and idioms matter.
\item
  Lazy-fetch when a package is used occasionally but source-level
  inspection may be needed.
\item
  Docs-first when the task targets public API, migration, conceptual
  usage, security guidance, or newly released features.
\item
  Index when repository or package scale exceeds grep-friendly size, or
  cross-file relationships dominate.
\item
  Skill when knowledge is procedural, stable, reusable, and safe to
  package.
\item
  Agentic-search when local source is available and the model/tooling
  can inspect selectively.
\item
  Pack when the agent lacks filesystem/search tools or the task is
  one-shot and offline.
\end{itemize}

\subsection{Source authority and conflict
resolution}\label{sec-authority}

A router should ask not only ``where can I get information?'' but
``which source is authoritative for this question?''

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.5000}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.5000}}@{}}
\caption{Source authority by question
type.}\label{tbl-authority}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Question
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Best authority
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Question
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Best authority
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
What version is installed? & Lockfile and package manager metadata \\
What public API is intended? & Official docs and public type surface \\
What behavior occurs? & Version-matched source and tests \\
What pattern does this repo prefer? & Local codebase and AGENTS.md \\
What should the agent never do? & AGENTS.md, security policy, human
instruction \\
What broke? & Failing test, log, trace, CI output \\
What changed between versions? & Changelog, release notes, migration
guide \\
What symbols reference this? & LSP, SCIP, source index, grep \\
\end{longtable}

Conflict resolution should preserve provenance. A useful agent-facing
message is not ``use zod docs''; it is ``this recommendation is based on
local package source resolved from \texttt{bun.lock}, not latest web
docs.'' That distinction prevents version skew and makes human override
possible.

\subsection{Provenance and context lockfiles}\label{sec-provenance}

The router should emit context plus provenance. Example bundle metadata
for source:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{context\_bundle}\KeywordTok{:}
\AttributeTok{  }\FunctionTok{source}\KeywordTok{:}\AttributeTok{ npm}
\AttributeTok{  }\FunctionTok{package}\KeywordTok{:}\AttributeTok{ effect}
\AttributeTok{  }\FunctionTok{version}\KeywordTok{:}\AttributeTok{ }\FloatTok{3.14.2}
\AttributeTok{  }\FunctionTok{route}\KeywordTok{:}\AttributeTok{ vendor}
\AttributeTok{  }\FunctionTok{resolved\_from}\KeywordTok{:}\AttributeTok{ bun.lock}
\AttributeTok{  }\FunctionTok{fetched\_at}\KeywordTok{:}\AttributeTok{ 2026{-}05{-}13}
\AttributeTok{  }\FunctionTok{trust\_class}\KeywordTok{:}\AttributeTok{ local{-}readonly{-}source}
\AttributeTok{  }\FunctionTok{mutable}\KeywordTok{:}\AttributeTok{ }\CharTok{false}
\AttributeTok{  }\FunctionTok{intended\_use}\KeywordTok{:}\AttributeTok{ reference{-}only}
\end{Highlighting}
\end{Shaded}

For docs:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{context\_bundle}\KeywordTok{:}
\AttributeTok{  }\FunctionTok{source}\KeywordTok{:}\AttributeTok{ context7}
\AttributeTok{  }\FunctionTok{package}\KeywordTok{:}\AttributeTok{ next}
\AttributeTok{  }\FunctionTok{version\_hint}\KeywordTok{:}\AttributeTok{ }\StringTok{"15.x"}
\AttributeTok{  }\FunctionTok{route}\KeywordTok{:}\AttributeTok{ docs}
\AttributeTok{  }\FunctionTok{fetched\_at}\KeywordTok{:}\AttributeTok{ 2026{-}05{-}13}
\AttributeTok{  }\FunctionTok{trust\_class}\KeywordTok{:}\AttributeTok{ remote{-}docs}
\AttributeTok{  }\FunctionTok{warning}\KeywordTok{:}\AttributeTok{ verify against lockfile and local source before editing}
\end{Highlighting}
\end{Shaded}

A \texttt{context.lock} file is the reproducibility bridge between
research and implementation. It records the strategy, source URL,
package version, commit, docs version, index version, skill version,
trust class, generated summaries, timestamps, and override reasons.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{[packages.effect]}
\DataTypeTok{strategy} \OperatorTok{=} \StringTok{"vendor"}
\DataTypeTok{version} \OperatorTok{=} \StringTok{"3.14.2"}
\DataTypeTok{source} \OperatorTok{=} \StringTok{"github:Effect{-}TS/effect"}
\DataTypeTok{commit} \OperatorTok{=} \StringTok{"abc123"}
\DataTypeTok{reason} \OperatorTok{=} \StringTok{"core{-}framework allowlist; imported in 42 files"}
\DataTypeTok{trust} \OperatorTok{=} \StringTok{"local{-}readonly"}

\KeywordTok{[packages.zod]}
\DataTypeTok{strategy} \OperatorTok{=} \StringTok{"docs{-}first"}
\DataTypeTok{version} \OperatorTok{=} \StringTok{"4.1.0"}
\DataTypeTok{docs} \OperatorTok{=} \StringTok{"context7:zod@4.1"}
\DataTypeTok{fallback} \OperatorTok{=} \StringTok{"fetch{-}source"}
\DataTypeTok{reason} \OperatorTok{=} \StringTok{"public API task; source size low; docs available"}
\DataTypeTok{trust} \OperatorTok{=} \StringTok{"remote{-}docs"}

\KeywordTok{[tools.mgrep]}
\DataTypeTok{enabled} \OperatorTok{=} \ConstantTok{false}
\DataTypeTok{reason} \OperatorTok{=} \StringTok{"privacy=strict{-}local"}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[]{@{}ll@{}}
\caption{Core fields for a context
lockfile.}\label{tbl-lockfields}\tabularnewline
\toprule\noalign{}
Lockfile field & Purpose \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
Lockfile field & Purpose \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\texttt{object\_id} & Stable identity for the routed artifact \\
\texttt{kind} & Context-object class \\
\texttt{source\_uri} & Upstream origin \\
\texttt{revision} & Commit, version, timestamp, or content hash \\
\texttt{authority} & Canonical, upstream, derived, advisory, or
unverified \\
\texttt{materialization} & Committed, cache, mount, live, or pack \\
\texttt{representation} & Raw, summary, metadata, index, or executable
tool \\
\texttt{privacy\_class} & Repo-safe, local-only, restricted, or
remote \\
\texttt{integrity} & Hash, signature, or provenance attestation \\
\texttt{phase\_policy} & Preferred routes by task phase \\
\texttt{refresh\_policy} & Pinned, manual refresh, TTL, or live \\
\texttt{transform\_chain} & Export, normalization, and summarization
steps \\
\end{longtable}

Provenance supports version fidelity, audit logs, privacy review,
benchmark reproducibility, and human overrides.

\subsection{Trust policy}\label{sec-security}

Security is central to routing, not an appendix. Every route has a
different attack surface.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.5000}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.5000}}@{}}
\caption{Route-specific security
risks.}\label{tbl-security}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Route
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Risk
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Route
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Risk
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Docs retrieval & Prompt injection, outdated examples, wrong version \\
MCP server & Tool overpermission, data exfiltration, remote execution \\
Skills & Executable scripts, hidden instructions, supply-chain
attacks \\
Vendored source & Malicious comments, enormous distractor surface \\
Repo pack & Secret exposure and uncontrolled redistribution \\
Cloud index & IP leakage and opaque retention \\
Agentic search & Command misuse and accidental broad reads \\
Generated summaries & Untraceable hallucinated compression \\
\end{longtable}

A practical router should support policy modes such as:

\begin{itemize}
\tightlist
\item
  \texttt{strict-local}: no remote retrieval or cloud index;
\item
  \texttt{local-executable-denied}: local files allowed, bundled scripts
  blocked;
\item
  \texttt{remote-docs-allowed}: remote docs allowed but quoted as
  untrusted;
\item
  \texttt{remote-code-denied}: no remote source fetch;
\item
  \texttt{cloud-index-allowed}: managed indexes permitted;
\item
  \texttt{scripts-require-confirmation}: skills and MCP tools cannot
  execute without confirmation;
\item
  \texttt{untrusted-context-quoted}: external context is evidence, not
  instruction.
\end{itemize}

The ethical case for an open router follows from these requirements:
transparent routing decisions, user override, local-first defaults,
provenance logs, auditable lockfiles, no silent cloud upload,
standards-based portability, and avoidance of vendor lock-in.

\subsection{Explainability requirements}\label{sec-explainability}

A useful routing explanation includes:

\begin{itemize}
\tightlist
\item
  selected strategy;
\item
  alternatives considered;
\item
  reason selected;
\item
  risks;
\item
  override command or config;
\item
  provenance;
\item
  expected cost.
\end{itemize}

Example:

\begin{quote}
Routed \texttt{effect} to \texttt{vendor} because it is imported in 47
files, appears in the core-framework allowlist, source size is below 25
MB, and privacy policy is \texttt{strict-local}. Alternatives:
docs-first rejected because the task requires idiomatic patterns; remote
index rejected by privacy policy. Override with
\texttt{ingraft\ route\ effect\ -\/-strategy\ fetch}.
\end{quote}

Explanations are not decoration. They are how users correct the policy.

\subsection{Worked routing examples}\label{sec-examples}

\textbf{Daily framework dependency.} A project uses Effect in 80 files.
The task is to implement a new service using local idioms. The router
vendors Effect source read-only, emits AGENTS.md guidance, and tells the
agent to inspect local examples before docs.

\textbf{Rare utility package.} The project uses \texttt{nanoid} in one
file and the task is unrelated. The router does nothing. If the agent
asks, it fetches source lazily.

\textbf{API migration.} A project upgrades Next.js from 14 to 15. The
router prioritizes version-specific migration docs and changelog, then
local usage search, then source fallback.

\textbf{Large monorepo.} The task touches one service in a 500K-file
monorepo. The router avoids full packing, queries structural code search
or SCIP/LSP, and passes only relevant files.

\textbf{Privacy-restricted enterprise repo.} Cloud retrieval is
disabled. The router uses local source, local docs, local symbolic
search, and blocks remote MCPs.

\section{Standards and Interoperability}\label{sec-standards}

Standards matter because routing should not be trapped inside one agent
vendor. The standards landscape is increasingly centered around AAIF
projects such as MCP, goose, and AGENTS.md, alongside adjacent standards
and conventions such as Agent Skills, ACP, A2A, llms.txt, and SCIP. It
would be wrong to imply that all of these have one governance story.

The Linux Foundation announced the Agentic AI Foundation on December 9,
2025, with founding project contributions including Anthropic's Model
Context Protocol, Block's goose, and OpenAI's AGENTS.md (Linux
Foundation 2025). The same announcement describes MCP as a protocol for
connecting models to tools, data, and applications, and AGENTS.md as a
project-guidance standard for coding agents. Agent Skills has its own
specification and progressive-disclosure model (Agent Skills Working
Group 2025; Anthropic 2025b). ACP, A2A, llms.txt, and SCIP have separate
origins and adoption paths (Zed Industries 2025; Google and Linux
Foundation 2025; Howard 2024; Sourcegraph 2024).

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.5000}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.5000}}@{}}
\caption{Standards and their routing
role.}\label{tbl-standards-role}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Standard or convention
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Router role
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Standard or convention
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Router role
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
AGENTS.md & Declares project-level routing policy and agent
instructions \\
MCP & Exposes router tools to multiple agents \\
Agent Skills & Packages strategy-specific procedures with progressive
disclosure \\
ACP & Lets editors invoke agents that call the router \\
A2A & Lets agents coordinate or delegate context work \\
SCIP/LSP & Supplies graph and symbol context \\
llms.txt & Provides docs-discovery hints, low authority unless
verified \\
\end{longtable}

The portable stack is therefore not ``one standard to rule them all.''
It is: AGENTS.md for durable project policy, MCP for tool access, Agent
Skills for procedural knowledge, and language/index standards for
structural retrieval.

%====================================================================
%  CUT: Evaluation Agenda. Field guides don't need eval agendas;
%  this material lives in the project repo's evals/ directory now.
%====================================================================
\iffalse
\section{Evaluation Agenda}\label{sec-eval}

The router thesis needs a router eval. The key research question is:

\begin{quote}
Does a task-aware context router outperform fixed context strategies
under equal model, budget, and tool conditions?
\end{quote}

\subsection{Strategies to compare}\label{sec-eval-strategies}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.5000}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.5000}}@{}}
\caption{Fixed strategies and router
condition.}\label{tbl-eval-strategies}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Strategy
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Description
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Strategy
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Description
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
No extra context & Baseline agent with repo only \\
Docs-only & Version-specific docs retrieval where available \\
Always-on docs index & AGENTS.md/CLAUDE.md points to local versioned
docs \\
Source vendoring & Dependency source in workspace \\
Lazy source fetch & Package source fetched on demand \\
Repo pack & Repomix-style snapshot \\
Semantic index & Embedding or code-search retrieval \\
Structural index & AST, LSP, SCIP, call/reference graph retrieval \\
Agentic search & Bash/grep/read loop \\
Skills/rules & Procedural guidance only \\
Router & Chooses among the above \\
\end{longtable}

\figureEvalLattice

\subsection{Task families}\label{sec-eval-tasks}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.5000}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.5000}}@{}}
\caption{Task families for router
evaluation.}\label{tbl-eval-tasks}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Task family
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Example
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Task family
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Example
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Unknown library use & Integrate unfamiliar package \\
Framework idiom task & Add route, middleware, schema, service, or
hook \\
API migration & Upgrade package version \\
Bugfix with stack trace & Failure crosses dependency boundary \\
Multi-file feature & Edit across local repo \\
Security-sensitive change & Avoid unsafe API or pattern \\
Performance/debugging task & Inspect implementation behavior \\
Test repair & Fix tests requiring project conventions \\
\end{longtable}

Stratification variables should include package popularity, source size,
docs quality, framework versus utility package, typed versus untyped
ecosystem, repo size, public API versus internals, package version
freshness, model class, and privacy setting.

\subsection{Metrics, ablations, and
falsification}\label{sec-eval-metrics}

\begin{longtable}[]{@{}ll@{}}
\caption{Router evaluation
metrics.}\label{tbl-eval-metrics}\tabularnewline
\toprule\noalign{}
Metric & Why it matters \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
Metric & Why it matters \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
pass@1 / pass@k & Task success \\
Tests passed & Functional correctness \\
Human review score & Maintainability and style \\
Tokens used & Context efficiency \\
Wall-clock latency & Practical usability \\
Tool calls & Loop complexity \\
Retrieval precision & Context quality \\
Provenance coverage & Auditability \\
Version match rate & Source/doc correctness \\
Regression rate & Safety \\
Cost per solved task & Operational relevance \\
\end{longtable}

Ablations should remove model awareness, privacy constraints, vendoring,
docs, graph/AST retrieval, stale-context detection, and
failure-triggered strategy switching. Stress tests should include stale
docs, distractor context, wrong package versions, and small versus
frontier models.

The thesis should be considered weakened if:

\begin{itemize}
\tightlist
\item
  one fixed strategy beats the router across most task/model/repo
  settings;
\item
  router explanations do not help users override bad choices;
\item
  routing overhead cancels out quality gains;
\item
  context strategy matters less than model choice once modern agents are
  used;
\item
  simpler agentic search dominates retrieval, indexing, and vendoring
  under realistic budgets.
\end{itemize}

Strong papers say what would falsify them. These are the falsification
tests.

\fi
%====================================================================
%  END CUT (Evaluation Agenda)
%====================================================================

\section{Where ingraft Fits}\label{sec-ingraft}

ingraft (Brunner 2026) is an open-source CLI that automates one specific
slot in the landscape: \textbf{vendoring third-party packages and
repositories into agent-ready reference directories}, with provenance,
update tooling, and \texttt{AGENTS.md} guidance baked in.

In the language of \S\ref{sec-strategies}, ingraft implements
Strategy~1 (vendoring) end-to-end and emits the Strategy~4 (instruct)
glue that keeps the vendored source read-only for humans while
remaining visible to agents. It does not try to be a full routing
runtime; the seven strategies coexist in any non-trivial workflow, and
the other six are best implemented by the tools surveyed in
\S\ref{sec-tools}.

\subsection*{What ingraft does today}
\addcontentsline{toc}{subsection}{What ingraft does today}

\begin{itemize}
\tightlist
\item
  Inspects the project's package manager state and lockfiles.
\item
  Adds chosen upstream repositories as committed subtrees under
  \texttt{vendor/}, with the same revision pin recorded in
  \texttt{.gitattributes} so PR diffs stay focused.
\item
  Supports four vendoring modes: \texttt{subtree} (committed source),
  \texttt{submodule} (gitlink), \texttt{clone-ignore} (local ignored
  clone), and \texttt{cache-link} (ignored symlink to a shared cache
  checkout).
\item
  Filters vendored repositories to omit media, generated directories,
  archives, fixtures, and oversized files when configured.
\item
  Generates \texttt{AGENTS.md} blocks listing what is vendored, the
  upstream URL and revision, and the rule that humans should not edit
  the vendored files.
\item
  Provides \texttt{list}, \texttt{add}, and \texttt{update}
  subcommands so the catalog stays maintainable.
\end{itemize}

\subsection*{When ingraft is the right pick}
\addcontentsline{toc}{subsection}{When ingraft is the right pick}

The fit matches Scenario~1 in \S\ref{sec-fit-scenario-effect}: a
project depends on a small set of idiom-heavy frameworks whose source
is more useful to coding agents than their human-facing docs, where
the team wants version-matched local source available to every agent
session without per-task fetches. Effect, Inertia.js, complex internal
TypeScript frameworks, and ``learn this library deeply'' onboarding
work are the natural domain.

\subsection*{When ingraft is the wrong pick}
\addcontentsline{toc}{subsection}{When ingraft is the wrong pick}

Vendoring is not the right strategy for migration tasks (Scenario~2),
multi-repo enterprise refactors (Scenario~4), one-shot Q\&A
(Scenario~6), or security-sensitive codebases where the constraint is
on-prem indexing rather than local source. For those, the tools and
strategy combinations in \S\ref{sec-fit} are better answers than
ingraft.

\subsection*{Project structure}
\addcontentsline{toc}{subsection}{Project structure}

A workspace using ingraft typically looks like:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{vendor/                         \# committed, read{-}only reference}
\NormalTok{  effect/                        \# subtree snapshot, version{-}pinned}
\NormalTok{  effect{-}smol/                   \# another subtree}

\NormalTok{AGENTS.md                        \# tells agents \textquotesingle vendor/ is reference\textquotesingle}
\NormalTok{.gitattributes                   \# marks vendor/ as vendored/generated}
\end{Highlighting}
\end{Shaded}

\noindent The full project layout and CLI reference are documented at
\url{https://ingraft.dev}.

%====================================================================
%  CUT: Threats to Validity, Predictions. Both were oriented to the
%  earlier "research agenda" framing. For a field guide, the relevant
%  caveats are folded into individual chapter prose.
%====================================================================
\iffalse
\section{Threats to Validity and Counterexamples}\label{sec-threats}

\textbf{Benchmark validity.} SWE-bench, LongCodeBench, CodeRAG-Bench,
RepoEval, CrossCodeEval, and vendor evals measure different task
distributions. Improvements on one may not transfer to another.

\textbf{Model drift.} Results from Claude 3.5, GPT-4o, Qwen2.5, Gemini
1.5, and other models may not extrapolate to newer systems. A router
should use model-aware calibration, not hardcoded thresholds.

\textbf{Tooling confounds.} Agent performance reflects model, scaffold,
tools, prompts, editing loop, test harness, retry policy, and human
oversight. It is hard to isolate context strategy alone.

\textbf{Vendor telemetry.} Cursor, Augment, Cognition, Greptile, and
similar numbers may be directionally useful but should not be treated as
independent evidence unless methodology is public.

\textbf{Publication bias.} Practitioner essays report successful
workflows more often than failed context strategies.

\textbf{Temporal instability.} Tool features, standards adoption,
pricing, security posture, and model context behavior change quickly.
Dated snapshots should be clearly labeled.

Routing may not help much in very small repos where the whole context
fits, tasks dominated by model reasoning rather than repo knowledge,
highly standardized frameworks where docs are enough, environments where
the agent already has excellent local tool use, security-constrained
settings where remote retrieval is forbidden and vendoring is
impossible, product-judgment tasks, repos with poor tests, languages
with weak structural tooling, generated or minified codebases, and
monorepos whose package boundaries do not match architectural
boundaries.

\section{Predictions}\label{sec-predictions}

The router thesis makes testable predictions:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Hybrid systems will outperform single-context systems under diverse
  task distributions.
\item
  The optimal route will vary by package centrality, not merely repo
  size.
\item
  Source vendoring will help most on idiom-heavy frameworks and least on
  stable utility libraries.
\item
  Docs-first retrieval will outperform source-first on migration tasks
  and public API usage.
\item
  Graph/AST retrieval will show larger gains as cross-file dependency
  depth increases.
\item
  Aggressive retrieval will hurt more often on distractor-heavy tasks
  than on tightly scoped tasks.
\item
  User-visible routing explanations will reduce repeated failure loops
  by making overrides possible.
\end{enumerate}

\fi
%====================================================================
%  END CUT (Threats and Predictions)
%====================================================================

\section{Conclusion}\label{sec-conclusion}

The 2026 context-engineering landscape is not a contest between
``vendor source'' and ``use RAG.'' It is a catalog --- seven strategies
that solve different parts of the same problem --- and a set of fit
rules that say which combination wins for which task.

Three operational principles fall out of the guide:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{No strategy dominates.} The seven win on different task
  classes, different scales, and different constraints. A team that
  picks one and uses it for everything is leaving accuracy and tokens
  on the table.
\item
  \textbf{Combinations are normal.} Almost every real workflow uses
  two or three strategies together: vendor + tool-mediated + instruct;
  docs index + symbol index + tool-mediated; pre-built index +
  tool-mediated + repo map. The interesting design choices are about
  the combination, not the primary.
\item
  \textbf{Constraints choose the route before quality does.} Privacy,
  freshness, latency, and trust often eliminate strategies before the
  accuracy comparison even starts. Picking a route means knowing
  which constraints bind first.
\end{enumerate}

The fit guide in \S\ref{sec-fit} is the artefact most readers will
return to. The strategies in \S\ref{sec-strategies} are the catalog
that explains what each row means. The tools in \S\ref{sec-tools} are
the implementations to reach for. The evidence in \S\ref{sec-evidence}
is the part to consult when the guide gets it wrong --- which it will,
because the field is moving and the May 2026 snapshot is dated by the
time it is read.

\begin{tcolorbox}[enhanced jigsaw, arc=.35mm, bottomrule=.15mm, bottomtitle=1mm, breakable, colback=white, colbacktitle=paper-callout-tip-color!10!white, colframe=paper-callout-tip-color-frame, coltitle=black, left=2mm, leftrule=.75mm, opacityback=0, opacitybacktitle=0.6, rightrule=.15mm, title=\textcolor{paper-callout-tip-color}{\faLightbulb}\hspace{0.5em}{Citation}, titlerule=0mm, toprule=.15mm, toptitle=1mm]

If you reference this work in academic or industry writing:

\begin{quote}
Brunner, G. \emph{Context Engineering for AI Coding Agents: A Field
Guide to Strategies, Tools, and Fit.} ingraft project, May 2026.
\url{https://ingraft.dev/whitepaper}
\end{quote}

Issues, corrections, and disagreements are welcome at
\href{https://github.com/gunta/ingraft/issues}{github.com/gunta/ingraft/issues}.

\end{tcolorbox}

\emph{This whitepaper is released under
\href{https://creativecommons.org/licenses/by/4.0/}{CC BY 4.0}. Quote
freely with attribution.}

\section{Appendix A: Representative Tool Snapshot, May
2026}\label{sec-tools-appendix}

This appendix preserves the tool-catalog value without letting the
catalog drive the main argument. It is a representative snapshot, not a
complete inventory.

\figureToolLandscapeCounts

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}@{}}
\caption{Representative May 2026 tool
snapshot.}\label{tbl-tool-snapshot}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Category
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Representative tools
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Routing lesson
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Category
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Representative tools
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Routing lesson
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
IDE-native context engines & Cursor, Windsurf, GitHub Copilot, Augment,
JetBrains AI/Junie, Cline, Zed, Roo Code, Continue, opencode, aider,
Sourcegraph Amp, Amazon Q, Devin, OpenHands, Factory Droid, Kiro, Google
Jules, Warp Agent Mode, Codebuff, Plandex & IDEs mix index, manual
mention, agentic search, and rules; the router should expose which layer
is active \\
Standalone code search/indexing & Sourcebot, Zoekt, Livegrep, Hound,
ast-grep, Comby, Bloop, Greptile, Glean Code Search, Unblocked, Tabby,
Serena MCP, CocoIndex, Indexify & Search wins when scale or cross-file
structure overwhelms live grep \\
Repo packaging/snapshots & Repomix, gitingest, code2prompt,
files-to-prompt, yek, ai-digest, gpt-repository-loader & Packs are
useful one-shot snapshots but weak on freshness and position control \\
MCP context servers & Context7, filesystem MCP, GitHub MCP, GitLab MCP,
DeepWiki MCP, fff, mgrep, Exa MCP, Tavily MCP, Perplexity MCP,
MCPfinder, Glama, Smithery, mcp.so, Memory Bank, Mem0, Letta, Mnemos,
Pieces LTM-2, OpenSrc & MCP turns context into tools; tool permissions
and provenance become router concerns \\
Embeddings/RAG components & voyage-code-3, Jina code embeddings, Jina v2
base code, mxbai-embed-large, CodeBERT, GraphCodeBERT, UniXcoder,
CodeT5+, StarCoder, CodeSage, ChromaDB, Qdrant, Pinecone, Weaviate,
Milvus, pgvector & Embeddings are substrates, not policies; quality
depends on selector, representation, and task \\
Multi-repo/org context & RepoSwarm, Backstage TechDocs, Roadie RAG,
Glean, Sourcegraph, DeepWiki, Unblocked & Org context requires ACLs,
provenance, and privacy-first routing \\
Adjacent safety/workflow & Qodo Merge, GitGuardian ggshield AI Hook,
MCP-omnisearch, GPTScript, Jujutsu, agentjj, jj-skill & Context routing
touches review, secrets, web search, shared prompts, and VCS
workflows \\
\end{longtable}

\section{Appendix B: Paper Matrix}\label{sec-paper-matrix}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedright\arraybackslash}p{(\linewidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedright\arraybackslash}p{(\linewidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedright\arraybackslash}p{(\linewidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedright\arraybackslash}p{(\linewidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedright\arraybackslash}p{(\linewidth - 10\tabcolsep) * \real{0.1667}}@{}}
\caption{Academic and technical works mapped to router
implications.}\label{tbl-paper-matrix}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Work
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Task
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Context type
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Main finding
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Router implication
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Evidence strength
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Work
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Task
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Context type
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Main finding
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Router implication
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Evidence strength
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
N. F. Liu et al. (2024) & QA & Long context & Position matters &
Prioritize placement and bundle size & Peer-reviewed journal \\
{Hsieh et al.} (2024) & Synthetic long-context & Multi-hop long context
& Effective length varies by model/task & Calibrate per model &
Peer-reviewed conference \\
Rando et al. (2025) & Code QA and repair & Long repo context & Long
context degrades unevenly on real code tasks & Avoid stuffing whole
repos; evaluate by task & Preprint/benchmark \\
Hong et al. (2025) & Long-context reliability & Distractors, semantic
needles, LongMemEval, repeated words & Increasing input length alone can
make outputs less reliable & Treat context rot as a first-class routing
cost & Technical report \\
Modarressi et al. (2025) & Long-context retrieval & Non-literal
needles & Literal matching overstates long-context capability & Prefer
semantic retrieval tests, not only lexical NIAH & Conference/preprint \\
Fu et al. (2025) & Missing-information detection & Omitted context &
Models struggle to identify absences even when retrieval looks easy &
Do not equate retrieval with comprehension & Conference/preprint \\
{Z. Z. Wang et al.} (2025) & Code generation & Retrieved docs/code &
Retrieval helps unevenly and retrievers miss useful contexts & Adaptive
retrieval, not fixed top-k & Peer-reviewed findings \\
Thakur et al. (2021) & Information retrieval & Heterogeneous retrieval
tasks & No retriever dominates zero-shot across domains & Route by task
distribution and retriever failure mode & Peer-reviewed benchmark \\
Khattab and Zaharia (2020) & Passage search & Late interaction &
Token-level matching improves retrieval while allowing precomputation &
Use late interaction when dense vectors are too lossy & Peer-reviewed
conference \\
Formal et al. (2021) & Information retrieval & Learned sparse
retrieval & Sparse expansion bridges lexical auditability and semantic
matching & Include sparse search in hybrid routing & Peer-reviewed
conference \\
Husain et al. (2019) & Code search & Natural-language to code & Code
search has a language gap distinct from generic text search & Evaluate
code retrieval separately from text retrieval & Benchmark/preprint \\
Li et al. (2024) & Code information retrieval & Multi-task code IR &
Code retrieval decomposes into different task families & Do not tune one
retriever for all code questions & Preprint/benchmark \\
Jiang et al. (2025) & Issue localization & Repository graph search &
Graph-guided search can outperform flatter localization routes & Add
graph search as an agentic fallback & Preprint \\
{``{U-NIAH}''} (2025) & Long-context/RAG & Synthetic needles/distractors
& RAG helps some models and hurts others & Model-aware routing &
Journal/preprint \\
F. Zhang et al. (2023) & Code completion & Iterative repo retrieval &
Repository context improves completion & Expand context iteratively &
Peer-reviewed conference \\
{Ding et al.} (2023) & Code completion & Cross-file context &
Single-file context is insufficient & Use repo-aware retrieval &
Peer-reviewed benchmark \\
{Liang et al.} (2024) & Code completion & Fused dual context & Context
quality and latency trade off & Rank, truncate, and score context &
Preprint \\
{Phan et al.} (2025) & Code completion & Semantic graph & Graph
retrieval outperforms flat baselines & Prefer structural retrieval when
possible & Workshop/preprint \\
{Y. (Sherry). Zhang et al.} (2025) & Code RAG & AST chunks & AST
chunking improves over line chunks & Structural chunking & Preprint \\
Brown (2026) & Practitioner source-vendoring report & Effect source tree
& Locally vendored source gives agents implementation structure, tests,
and examples that docs and web fetches omit & Treat source vendoring as
a candidate route for high-idiom, high-centrality packages &
Practitioner/vendor evidence \\
Gao (2026a) & Framework-agent evals & AGENTS.md docs index versus skills
& Always-on compressed docs index outperformed on-demand skill use in
their Next.js suite & Evaluate selector/timing, not only content quality
& Vendor engineering eval \\
Shi et al. (2023) & Reasoning & Irrelevant context & Distractors reduce
performance & Penalize distractor risk & Peer-reviewed conference \\
\end{longtable}

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\protect\phantomsection\label{refs}
\begin{CSLReferences}{1}{1}
\bibitem[\citeproctext]{ref-agentskills2025spec}
Agent Skills Working Group. 2025. {``Agent Skills Specification.''}
\url{https://agentskills.io/specification}.

\bibitem[\citeproctext]{ref-anthropic2025contexteng}
Anthropic. 2025a. {``Effective Context Engineering for {AI} Agents.''}
September 29.
\url{https://www.anthropic.com/engineering/effective-context-engineering-for-ai-agents}.

\bibitem[\citeproctext]{ref-anthropic2025skills}
Anthropic. 2025b. {``Equipping Agents for the Real World with Agent
Skills.''} October 16.
\url{https://www.anthropic.com/engineering/equipping-agents-for-the-real-world-with-agent-skills}.

\bibitem[\citeproctext]{ref-asai2024selfrag}
Asai, Akari, Zeqiu Wu, Yizhong Wang, Avirup Sil, and Hannaneh
Hajishirzi. 2024. {``Self-{RAG}: Learning to Retrieve, Generate, and
Critique Through Self-Reflection.''} \emph{ICLR 2024}.
\url{https://arxiv.org/abs/2310.11511}.

\bibitem[\citeproctext]{ref-cline2025nondex}
Baumann, Nick. 2025. {``Why {Cline} Doesn't Index Your Codebase (and Why
That's a Good Thing).''} May 27.
\url{https://cline.bot/blog/why-cline-doesnt-index-your-codebase-and-why-thats-a-good-thing}.

\bibitem[\citeproctext]{ref-brown2026effect}
Brown, Maxwell. 2026. {``The One Weird Git Trick That Makes Coding
Agents More {Effect}-Ive.''} May 11.
\url{https://effect.website/blog/the-one-weird-git-trick-that-makes-coding-agents-more-effect-ive/}.

\bibitem[\citeproctext]{ref-ingraft2026}
Brunner, Gunther. 2026. {``{ingraft}: The Context Router for {AI} Coding
Agents.''} \url{https://ingraft.dev/}.

\bibitem[\citeproctext]{ref-cursor2026indexing}
Cursor. 2026. {``Securely Indexing Large Codebases.''} January 27.
\url{https://cursor.com/blog/secure-codebase-indexing}.

\bibitem[\citeproctext]{ref-ding2023crosscodeeval}
{Ding, Yangruibo et al.} 2023. {``{CrossCodeEval}: A Diverse and
Multilingual Benchmark for Cross-File Code Completion.''} \emph{NeurIPS
2023 Datasets and Benchmarks Track}.
\url{https://arxiv.org/abs/2310.11248}.

\bibitem[\citeproctext]{ref-ding2024cocomic}
Ding, Yangruibo, Zijian Wang, Wasi Uddin Ahmad, et al. 2024.
{``{CoCoMIC}: Code Completion by Jointly Modeling in-File and Cross-File
Context.''} \emph{Proceedings of LREC-COLING 2024}.
\url{https://arxiv.org/abs/2212.10007}.

\bibitem[\citeproctext]{ref-vercel2026agentsmd}
Gao, Jude. 2026a. {``{AGENTS.md} Outperforms Skills in Our Agent
Evals.''} January 27.
\url{https://vercel.com/blog/agents-md-outperforms-skills-in-our-agent-evals}.

\bibitem[\citeproctext]{ref-vercel2026nextagentspr}
Gao, Jude. 2026b. {``Feat(next-Codemod): Add Agents-Md Command for {AI}
Coding Agents.''} January 24.
\url{https://github.com/vercel/next.js/pull/88961}.

\bibitem[\citeproctext]{ref-gauthier2023repomap}
Gauthier, Paul. 2023. {``Building a Better Repository Map with
Tree-Sitter.''} October 22.
\url{https://aider.chat/2023/10/22/repomap.html}.

\bibitem[\citeproctext]{ref-a2a2025spec}
Google, and Linux Foundation. 2025. {``Agent-to-Agent Protocol
({A2A}).''} \url{https://a2a-protocol.org/latest/}.

\bibitem[\citeproctext]{ref-llmstxt2024spec}
Howard, Jeremy. 2024. {``The {llms.txt} Specification.''} September 3.
\url{https://llmstxt.org/}.

\bibitem[\citeproctext]{ref-hsieh2024ruler}
{Hsieh, Cheng-Ping et al.} 2024. {``{RULER}: What's the Real Context
Size of Your Long-Context Language Models?''} \emph{COLM 2024}.
\url{https://arxiv.org/abs/2404.06654}.

\bibitem[\citeproctext]{ref-humanlayer2025claudemd}
HumanLayer (Kyle). 2025. {``Writing a Good {CLAUDE.md}.''} November 25.
\url{https://www.humanlayer.dev/blog/writing-a-good-claude-md}.

\bibitem[\citeproctext]{ref-liang2024repofuse}
{Liang, Ming, Xiaoheng Xie, Gehao Zhang, Xunjin Zheng, Peng Di, et al.}
2024. \emph{{REPOFUSE}: Repository-Level Code Completion with Fused Dual
Context}. \url{https://arxiv.org/abs/2402.14323}.

\bibitem[\citeproctext]{ref-aaif2025foundation}
Linux Foundation. 2025. {``{Linux Foundation} Announces the Formation of
the {Agentic AI Foundation}.''} December.
\url{https://www.linuxfoundation.org/press/linux-foundation-announces-the-formation-of-the-agentic-ai-foundation}.

\bibitem[\citeproctext]{ref-liu2024lost}
Liu, Nelson F., Kevin Lin, John Hewitt, et al. 2024. {``Lost in the
Middle: How Language Models Use Long Contexts.''} \emph{Transactions of
the Association for Computational Linguistics}.
\url{https://arxiv.org/abs/2307.03172}.

\bibitem[\citeproctext]{ref-liu2024graphcoder}
{Liu, Wei et al.} 2024. {``{GraphCoder}: Code Context Graph-Based
Retrieval for Repository-Level Completion.''} \emph{ASE 2024}.
\url{https://arxiv.org/abs/2406.07003}.

\bibitem[\citeproctext]{ref-phan2025repohyper}
{Phan, Huy N. et al.} 2025. {``{RepoHyper}: Search-Expand-Refine on
Semantic Graphs for Repository-Level Code Completion.''} \emph{FORGE
2025}. \url{https://arxiv.org/abs/2403.06095}.

\bibitem[\citeproctext]{ref-longcodebench2025}
Rando, Stefano, Luca Romani, Alessio Sampieri, et al. 2025.
\emph{{LongCodeBench}: Evaluating Coding LLMs at 1M Context Windows}.
arXiv preprint. \url{https://arxiv.org/abs/2505.07897}.

\bibitem[\citeproctext]{ref-shi2023distracted}
Shi, Freda, Xinyun Chen, Kanishka Misra, et al. 2023. {``Large Language
Models Can Be Easily Distracted by Irrelevant Context.''} \emph{ICML
2023}. \url{https://arxiv.org/abs/2302.00093}.

\bibitem[\citeproctext]{ref-shrivastava2023repofusion}
Shrivastava, Disha, Denis Kocetkov, Harm de Vries, Dzmitry Bahdanau, and
Torsten Scholak. 2023. \emph{{RepoFusion}: Training Code Models to
Understand Your Repository}. arXiv preprint.
\url{https://arxiv.org/abs/2306.10998}.

\bibitem[\citeproctext]{ref-sourcegraph2024scip}
Sourcegraph. 2024. {``{SCIP} --- {SourceCode Indexing Protocol}.''}
\url{https://github.com/sourcegraph/scip}.

\bibitem[\citeproctext]{ref-sourcegraph2025amp}
Sourcegraph. 2025. {``Why {Sourcegraph} and {Amp} Are Becoming
Independent Companies.''} December.
\url{https://sourcegraph.com/blog/why-sourcegraph-and-amp-are-becoming-independent-companies}.

\bibitem[\citeproctext]{ref-uniah2025}
{``{U-NIAH}: Unified {RAG} and {LLM} Evaluation for Long Context
Needle-in-a-Haystack.''} 2025. \emph{ACM Transactions on Information
Systems}. \url{https://arxiv.org/abs/2503.00353}.

\bibitem[\citeproctext]{ref-vercel2026nextevalsoss}
Vercel. 2026. {``{next-evals-oss}: Agent Evaluations for {Next.js}
Coding Tasks.''} \url{https://github.com/vercel/next-evals-oss}.

\bibitem[\citeproctext]{ref-latentspace2025claudecode}
Wang, Shawn (swyx), and Alessio Trotta. 2025. {``{Latent Space} Podcast:
{Claude Code} with Boris Cherny and Catherine Wu.''} May 7.
\url{https://www.latent.space/p/claude-code}.

\bibitem[\citeproctext]{ref-wang2025coderagbench}
{Wang, Zora Zhiruo et al.} 2025. {``{CodeRAG-Bench}: Can Retrieval
Augment Code Generation?''} \emph{NAACL Findings 2025}.
\url{https://arxiv.org/abs/2406.14497}.

\bibitem[\citeproctext]{ref-zed2025acp}
Zed Industries. 2025. {``Agent Client Protocol ({ACP}).''}
\url{https://zed.dev/acp}.

\bibitem[\citeproctext]{ref-zhang2023repocoder}
Zhang, Fengji, Bei Chen, Yue Zhang, et al. 2023. {``{RepoCoder}:
Repository-Level Code Completion Through Iterative Retrieval and
Generation.''} \emph{Proceedings of EMNLP 2023}.
\url{https://arxiv.org/abs/2303.12570}.

\bibitem[\citeproctext]{ref-zhang2025cast}
{Zhang, Yilin (Sherry) et al.} 2025. \emph{{cAST}: Enhancing Code RAG
with Structural Chunking via Abstract Syntax Tree}.
\url{https://arxiv.org/abs/2506.15655}.

\bibitem[\citeproctext]{ref-chen2024bgem3}
Chen, Jianlv, Shitao Xiao, Peitian Zhang, Kun Luo, Defu Lian, and Zheng
Liu. 2024. \emph{{BGE M3}-Embedding: Multi-Lingual,
Multi-Functionality, Multi-Granularity Text Embeddings Through
Self-Knowledge Distillation}. arXiv preprint.
\url{https://arxiv.org/abs/2402.03216}.

\bibitem[\citeproctext]{ref-formal2021splade}
Formal, Thibault, Benjamin Piwowarski, and Stéphane Clinchant. 2021.
{``{SPLADE}: Sparse Lexical and Expansion Model for First Stage
Ranking.''} \emph{SIGIR 2021}. \url{https://arxiv.org/abs/2107.05720}.

\bibitem[\citeproctext]{ref-fu2025absencebench}
Fu, Harvey Yiyun, Aryan Shrivastava, Jared Moore, Peter West, Chenhao
Tan, and Ari Holtzman. 2025. {``AbsenceBench: Language Models Can't Tell
What's Missing.''} \emph{NeurIPS 2025 Datasets and Benchmarks Track}.
\url{https://arxiv.org/abs/2506.11440}.

\bibitem[\citeproctext]{ref-hong2025contextrot}
Hong, Kelly, Anton Troynikov, and Jeff Huber. 2025. {``Context Rot: How
Increasing Input Tokens Impacts LLM Performance.''} Chroma Technical
Report. \url{https://trychroma.com/research/context-rot}.

\bibitem[\citeproctext]{ref-husain2019codesearchnet}
Husain, Hamel, Ho-Hsiang Wu, Tiferet Gazit, Miltiadis Allamanis, and
Marc Brockschmidt. 2019. \emph{CodeSearchNet Challenge: Evaluating the
State of Semantic Code Search}. arXiv preprint.
\url{https://arxiv.org/abs/1909.09436}.

\bibitem[\citeproctext]{ref-jiang2025cosil}
Jiang, Zhonghao, Xiaoxue Ren, Meng Yan, Wei Jiang, Yong Li, and Zhongxin
Liu. 2025. \emph{{CoSIL}: Software Issue Localization via LLM-Driven
Code Repository Graph Searching}. arXiv preprint.
\url{https://arxiv.org/abs/2503.22424}.

\bibitem[\citeproctext]{ref-khattab2020colbert}
Khattab, Omar, and Matei Zaharia. 2020. {``{ColBERT}: Efficient and
Effective Passage Search via Contextualized Late Interaction over
BERT.''} \emph{Proceedings of SIGIR 2020}, 39--48.
\url{https://doi.org/10.1145/3397271.3401075}.

\bibitem[\citeproctext]{ref-lee2024mxbaiembed}
Lee, Sean, Aamir Shakir, Darius Koenig, and Julius Lipp. 2024a. {``Open
Source Strikes Bread -- New Fluffy Embedding Model.''} Mixedbread.
\url{https://www.mixedbread.com/blog/mxbai-embed-large-v1}.

\bibitem[\citeproctext]{ref-lee2024mxbaicolbert}
Lee, Sean, Aamir Shakir, Darius Koenig, and Julius Lipp. 2024b.
{``ColBERTus Maximus -- Introducing mxbai-colbert-large-v1.''}
Mixedbread. \url{https://www.mixedbread.com/blog/mxbai-colbert-large-v1}.

\bibitem[\citeproctext]{ref-li2024coir}
Li, Xiangyang, Kuicai Dong, Yi Quan Lee, Wei Xia, Yichun Yin, Hao Zhang,
Yong Liu, Yasheng Wang, and Ruiming Tang. 2024. \emph{{CoIR}: A
Comprehensive Benchmark for Code Information Retrieval Models}. arXiv
preprint. \url{https://arxiv.org/abs/2407.02883}.

\bibitem[\citeproctext]{ref-mixedbread2026mgrep}
Mixedbread AI. 2026. {``mgrep: Semantic Search for Agents.''}
\url{https://www.mgrep.dev/}.

\bibitem[\citeproctext]{ref-modarressi2025nolima}
Modarressi, Ali, Hanieh Deilamsalehy, Franck Dernoncourt, Trung Bui,
Ryan A. Rossi, Seunghyun Yoon, and Hinrich Schütze. 2025.
{``NoLiMa: Long-Context Evaluation Beyond Literal Matching.''}
\emph{International Conference on Machine Learning}.
\url{https://arxiv.org/abs/2502.05167}.

\bibitem[\citeproctext]{ref-mohsenimofidi2025contextoss}
Mohsenimofidi, Seyedmoein, Matthias Galster, Christoph Treude, and
Sebastian Baltes. 2025. \emph{Context Engineering for AI Agents in
Open-Source Software}. arXiv preprint.
\url{https://arxiv.org/abs/2510.21413}.

\bibitem[\citeproctext]{ref-muennighoff2023mteb}
Muennighoff, Niklas, Nouamane Tazi, Loic Magne, and Nils Reimers. 2023.
{``{MTEB}: Massive Text Embedding Benchmark.''} \emph{Proceedings of
EACL 2023}. \url{https://arxiv.org/abs/2210.07316}.

\bibitem[\citeproctext]{ref-robertson2009bm25}
Robertson, Stephen, and Hugo Zaragoza. 2009. {``The Probabilistic
Relevance Framework: {BM25} and Beyond.''} \emph{Foundations and Trends
in Information Retrieval} 3 (4): 333--389.
\url{https://doi.org/10.1561/1500000019}.

\bibitem[\citeproctext]{ref-santhanam2022colbertv2}
Santhanam, Keshav, Omar Khattab, Jon Saad-Falcon, Christopher Potts, and
Matei Zaharia. 2022a. {``{ColBERTv2}: Effective and Efficient Retrieval
via Lightweight Late Interaction.''} \emph{NAACL 2022}.
\url{https://aclanthology.org/2022.naacl-main.272/}.

\bibitem[\citeproctext]{ref-santhanam2022plaid}
Santhanam, Keshav, Omar Khattab, Christopher Potts, and Matei Zaharia.
2022b. \emph{{PLAID}: An Efficient Engine for Late Interaction
Retrieval}. arXiv preprint. \url{https://arxiv.org/abs/2205.09707}.

\bibitem[\citeproctext]{ref-thakur2021beir}
Thakur, Nandan, Nils Reimers, Andreas Rücklé, Abhishek Srivastava, and
Iryna Gurevych. 2021. {``{BEIR}: A Heterogeneous Benchmark for
Zero-Shot Evaluation of Information Retrieval Models.''} \emph{NeurIPS
2021 Datasets and Benchmarks Track}.
\url{https://openreview.net/forum?id=wCu6T5xFjeJ}.

\bibitem[\citeproctext]{ref-thakur2025freshstack}
Thakur, Nandan, Jimmy Lin, Sam Havens, Michael Carbin, Omar Khattab, and
Andrew Drozdov. 2025. {``FreshStack: Building Realistic Benchmarks for
Evaluating Retrieval on Technical Documents.''} \emph{NeurIPS 2025
Datasets and Benchmarks Track}. \url{https://arxiv.org/abs/2504.13128}.

\end{CSLReferences}




\end{document}
